### 2014-06-20 README.Rmd
Added README.md based on R Markdown file.

---
title: "README"
output: html_document
---

SpatialEpi: Methods and Data for Spatial Epidemiology
===========================================
We load the data and convert the coordinate system from latitude/longitude to a 
grid-based system.  
```{r}
require(SpatialEpi)
data(NYleukemia)
map <- NYleukemia$spatial.polygon
centroids <- latlong2grid(NYleukemia$geo[, 2:3])
population <- NYleukemia$data$population
cases <- NYleukemia$data$cases
```

There were 4 census tracts that were completely surrounded by another census tract (an enclave like Lesotho).  We merge them into their surrounding census tracts and modify the population/case counts and SpatialPolygons object accordingly. 
```{r}
remove <- NYleukemia$surrounded
add <- NYleukemia$surrounding
population[add] <- population[add] + population[remove]
population <- population[-remove]
cases[add] <- cases[add] + cases[remove]
cases <- cases[-remove]
centroids <- centroids[-remove, ]
map <- SpatialPolygons(map@polygons[-remove], proj4string=CRS("+proj=longlat"))
```

We plot the incidence of leukemia for each census tract.  
```{r, fig.height=8, fig.width=8}
plotmap(cases/population, map, log=TRUE, nclr=5)
points(grid2latlong(centroids), pch=4)
```

### 2014-06-20 README.Rmd
Added plot
More toying with README plot.

@@
 ---
-title: "README"
-output: html_document
+output:
+  html_document: default
 ---
 
-SpatialEpi: Methods and Data for Spatial Epidemiology
+SpatialEpi
 ===========================================
+Methods and Data for Spatial Epidemiology
+-------------------------------------------
 We load the data and convert the coordinate system from latitude/longitude to a 
@@
 We plot the incidence of leukemia for each census tract.  
-```{r, fig.height=8, fig.width=8}
+```{r, fig.height=8, fig.width=8, fig.path='./figure/'}
 plotmap(cases/population, map, log=TRUE, nclr=5)

### 2014-08-13 README.Rmd
Updated README
Updated README.md file.

@@
 output:
-  html_document: default
+  html_document:
+    keep_md: yes
 ---
@@
 ```{r}
-require(SpatialEpi)
+library(SpatialEpi)
 data(NYleukemia)
-map <- NYleukemia$spatial.polygon
+sp.obj <- NYleukemia$spatial.polygon
 centroids <- latlong2grid(NYleukemia$geo[, 2:3])
@@
 
-There were 4 census tracts that were completely surrounded by another census tract (an enclave like Lesotho).
-We merge them into their surrounding census tracts and modify the population/case counts and SpatialPolygons object accordingly.
-```{r}
-remove <- NYleukemia$surrounded
-add <- NYleukemia$surrounding
-population[add] <- population[add] + population[remove]
-population <- population[-remove]
-cases[add] <- cases[add] + cases[remove]
-cases <- cases[-remove]
-centroids <- centroids[-remove, ]
-map <- SpatialPolygons(map@polygons[-remove], proj4string=CRS("+proj=longlat"))
+We plot the incidence of leukemia for each census tract.  
+```{r, fig.height=8, fig.width=8, fig.path='./figure/'}
+plotmap(cases/population, sp.obj, log=TRUE, nclr=5)
+points(grid2latlong(centroids), pch=4)
 ```
 
-We plot the incidence of leukemia for each census tract.  
-```{r, fig.height=8, fig.width=8, fig.path='./figure/'}
-plotmap(cases/population, map, log=TRUE, nclr=5)
-points(grid2latlong(centroids), pch=4)
+We run the Bayesian Cluster Detection method from Wakefield and Kim (2013)
+```{r, fig.height=8, fig.width=8, fig.path='./figure/', cache=TRUE}
+y <- cases
+E <- expected(population, cases, 1)
+max.prop <- 0.15
+shape <- c(2976.3, 2.31)
+rate <- c(2977.3, 1.31)
+J <- 7
+pi0 <- 0.95
+n.sim.lambda <- 10^4
+n.sim.prior <- 10^5
+n.sim.post <- 10^5
+
+# Compute output
+output <- bayes_cluster(y, E, population, sp.obj, centroids, max.prop, 
+ shape, rate, J, pi0, n.sim.lambda, n.sim.prior, n.sim.post)
+plotmap(output$post.map$high.area, sp.obj)
 ```

### 2015-09-30 R/SpatialEpi.R
Start massive re-org: Following Hadley Wickham's R Packages book, creating Vignette, incorporation roxygen2 & devtools tools.

@@ -1,0 +1,12 @@
+#' SpatialEpi: A package for methods and data for spatial epidemiology.
+#'
+#' The SpatialEpi package provides three categories of important functions:
+#' SpatialEpi, bar and baz.
+#' 
+#' @section SpatialEpi functions:
+#' The SpatialEpi functions ...
+#'
+#' @docType package
+#' @name SpatialEpi
+NULL
+#> NULL

### 2015-09-30 R/circle.R

@@ -1,17 +1,37 @@
-circle <-
-function(geo, cluster.center, cluster.end){
-
-# Compute interpoint distance
-distance <- dist(as.matrix(geo), upper=TRUE, diag=TRUE)
-distance <- as.matrix(distance)[cluster.center, cluster.end]
-
-# For drawing radius of cluster on map
-polar <- seq(0, 2*pi, length=1000)
-        
-cluster.radius <- data.frame(cbind(
-  x=distance*cos(polar)+geo$x[cluster.center],
-  y=distance*sin(polar)+geo$y[cluster.center]
-  ))
-
-return(cluster.radius)
+#' Compute cartesian coordinates of a cluster center and radius
+#' 
+#' This function is used for plotting purposes
+#' 
+#' @param geo A \code{n x 2} table of the x-coordinate and y-coordinates of the 
+#'   centroids of each area
+#' @param cluster.center The area index (an integer between \code{1} and
+#'   \code{n}) indicating the center of the circle
+#' @param cluster.end The area index (an integer between \code{1} and \code{n})
+#'   indicating the area at the end of the circle
+#'   
+#' @return cluster.radius A data frame that you can plot
+#'   
+#' @author Albert Y. Kim
+#' @seealso \code{\link{zones}}
+#'   
+#' @examples
+#' data(pennLC)
+#' geo <- pennLC$geo[,2:3]
+#' plot(geo,type='n')
+#' text(geo,labels=1:nrow(geo))
+#' lines( circle(geo, 23, 46), col = "red" ) 
+#'   
+circle <- function(geo, cluster.center, cluster.end) {
+  # Compute interpoint distance
+  distance <- dist(as.matrix(geo), upper = TRUE, diag = TRUE)
+  distance <- as.matrix(distance)[cluster.center, cluster.end]
+  
+  # For drawing radius of cluster on map
+  polar <- seq(0, 2 * pi, length = 1000)
+  
+  cluster.radius <- data.frame(cbind(x = distance * cos(polar) + geo$x[cluster.center], 
+    y = distance * sin(polar) + geo$y[cluster.center]))
+  
+  return(cluster.radius)
 }
+

### 2015-09-30 R/eBayes.R

@@ -1,5 +1,4 @@
-eBayes <-
-function(Y, E, Xmat=NULL){
+eBayes <- function(Y, E, Xmat=NULL){
 	# Check for covariates
 	if (is.null(Xmat)) {
 		mod <- glm.nb(Y ~ 1+offset(log(E)), link=log)

### 2015-09-30 R/plot_crayons.R

@@ -1,0 +1,65 @@
+#' Illustration of crayon colors
+#'
+#' Creates a plot of the crayon colors in \code{\link{brocolors}}
+#'
+#' @param method2order method to order colors (\code{"hsv"} or \code{"cluster"})
+#' @param cex character expansion for the text
+#' @param mar margin paramaters; vector of length 4 (see \code{\link[graphics]{par}})
+#'
+#' @return None
+#'
+#' @author Karl W Broman, \email{kbroman@@biostat.wisc.edu}
+#' @references \url{http://en.wikipedia.org/wiki/List_of_Crayola_crayon_colors}
+#' @seealso \code{\link{brocolors}}
+#' @keywords hplot
+#'
+#' @examples
+#' plot_crayons()
+#'
+#' @export
+#' @importFrom grDevices rgb2hsv
+#' @importFrom graphics par plot rect text
+#'
+plot_crayons <-
+  function(method2order=c("hsv", "cluster"), cex=0.6, mar=rep(0.1, 4))
+  {
+    method2order <- match.arg(method2order)
+    
+    crayons <- brocolors("crayons")
+    
+    # get rgb
+    colval <- col2rgb(crayons)
+    
+    if(method2order == "hsv") {
+      # convert to hsv
+      colval <- t(rgb2hsv(colval))
+      
+      # order the colors; first two lines are to get black/gray/silver/white first
+      ord <- order(names(crayons)!="Black", names(crayons)!="Gray",
+        names(crayons)!="Silver", names(crayons)!="White",
+        colval[,1], colval[,2], colval[,3])
+      
+    } else {
+      ord <- hclust(dist(t(colval)))$ord
+    }
+    
+    oldmar <- par("mar")
+    on.exit(par(mar=oldmar))
+    
+    par(mar=mar)
+    x <- (1:7)-1
+    y <- (1:19)-1
+    x <- rep(x, each=19)
+    y <- rep(y, 7)
+    
+    plot(0, 0, type="n", xlab="", ylab="", xaxs="i", yaxs="i",
+      xlim=c(0, max(x)+1), ylim=c(max(y)+0.5, -0.5),
+      xaxt="n", yaxt="n")
+    
+    dx <- 0.2
+    dy <- 0.4
+    rect(x+dx/4, y-dy, x+dx, y+dy, border="black",
+      col=crayons[ord])
+    
+    text(x+dx*1.2, y, names(crayons)[ord], cex=cex, adj=c(0, 0.5))
+  }

### 2015-10-02 R/RcppExports.R
Phase 1 of rebuild.

@@ -1,71 +1,0 @@
-# This file was generated by Rcpp::compileAttributes
-# Generator token: 10BE3573-1514-4C36-9D1C-5A225CD40393
-
-normalize <- function(x) {
-    .Call('SpatialEpi_normalize', PACKAGE = 'SpatialEpi', x)
-}
-
-NumericVectorEquality <- function(x, y) {
-    .Call('SpatialEpi_NumericVectorEquality', PACKAGE = 'SpatialEpi', x, y)
-}
-
-ldnbinom <- function(y, E, a, b) {
-    .Call('SpatialEpi_ldnbinom', PACKAGE = 'SpatialEpi', y, E, a, b)
-}
-
-ldmultinom <- function(x, prob) {
-    .Call('SpatialEpi_ldmultinom', PACKAGE = 'SpatialEpi', x, prob)
-}
-
-coeff <- function(y_vector, E_vector, a_values, b_values, cluster_list) {
-    .Call('SpatialEpi_coeff', PACKAGE = 'SpatialEpi', y_vector, E_vector, a_values, b_values, cluster_list)
-}
-
-ProbSampleReplace <- function(prob) {
-    .Call('SpatialEpi_ProbSampleReplace', PACKAGE = 'SpatialEpi', prob)
-}
-
-check_overlap <- function(config, overlap) {
-    .Call('SpatialEpi_check_overlap', PACKAGE = 'SpatialEpi', config, overlap)
-}
-
-clean_moves_matrix <- function(theta, moves, n_zones) {
-    .Call('SpatialEpi_clean_moves_matrix', PACKAGE = 'SpatialEpi', theta, moves, n_zones)
-}
-
-return_death_moves <- function(theta) {
-    .Call('SpatialEpi_return_death_moves', PACKAGE = 'SpatialEpi', theta)
-}
-
-return_birth_moves <- function(theta, overlap) {
-    .Call('SpatialEpi_return_birth_moves', PACKAGE = 'SpatialEpi', theta, overlap)
-}
-
-return_local_moves <- function(theta, overlap, cluster_coords) {
-    .Call('SpatialEpi_return_local_moves', PACKAGE = 'SpatialEpi', theta, overlap, cluster_coords)
-}
-
-MCMC_simulation <- function(n_sim, pattern, theta_init, overlap, cluster_coords, p_moves_orig, J, lkhd_z, lambda) {
[…]

### 2015-10-02 R/SpatialEpi.R

@@ -1,12 +1,13 @@
-#' SpatialEpi: A package for methods and data for spatial epidemiology.
+#' SpatialEpi: Data and methods for spatial epidemiology.
 #'
-#' The SpatialEpi package provides three categories of important functions:
-#' SpatialEpi, bar and baz.
+#' This package provides...
 #' 
-#' @section SpatialEpi functions:
-#' The SpatialEpi functions ...
+#' @section Foo functions:
+#' The foo functions ...
 #'
 #' @docType package
+#' @import dplyr
+#' @import sp
 #' @name SpatialEpi
 NULL
 #> NULL

### 2015-10-02 R/bayes_cluster.R

@@ -1,108 +1,0 @@
-bayes_cluster <-
-function(y, E, population, sp.obj, centroids, max.prop, shape, rate, J, pi0,
-         n.sim.lambda, n.sim.prior, n.sim.post, burnin.prop = 0.1,
-         theta.init = vector(mode="numeric", length=0)
-){
-print(paste("Algorithm started on:", date()))
-  
-# MCMC parameters
-pattern <- c(0, 1)
-p.moves <- normalize(c(1, 1, 1, 1, 1))
-  
-# Geographic info and create geographical objects to use
-geo.objects <- create_geo_objects(max.prop, population, centroids, sp.obj)
-overlap <- geo.objects$overlap
-cluster.coords <- geo.objects$cluster.coords    
-n <- length(E)
-n.zones <- nrow(cluster.coords)
-print(paste("Geographic objects creation complete on:", date()))
-
-# Set cutoffs of the relative risk to determine low and high risk
-RR.range <- seq(0.7, 1/0.7, length=100000)
-pN <- dgamma(RR.range, shape[1], rate[1])
-pW <- dgamma(RR.range, shape[2], rate[2])
-cutoffs <- list(
-  low=RR.range[min(which(pN/pW>1))],
-  high=RR.range[max(which(pN/pW>1))]
-)
-
-
-#-------------------------------------------------------------------------------
-# Obtain values of lambda using importance sampling
-#-------------------------------------------------------------------------------
-# Set uniform prior on single zones.
-prior.z <- rep(1/n.zones, n.zones)
-log_prior.z <- log(prior.z) - log(sum(prior.z))
-
-# Estimate q and lambda
-results <- estimate_lambda(n.sim.lambda, J, prior.z, overlap, pi0)
-lambda <- results$lambda
-prior.j <- results$prior.j
-print(paste("Importance sampling of lambda complete on:", date()))
-
-
-#-------------------------------------------------------------------------------
-# Obtain prior map
-#-------------------------------------------------------------------------------
-# Generate MCMC samples from prior
-prior.chain <- MCMC_simulation(n.sim.prior, pattern, theta.init, overlap, 
-                               cluster.coords, p.moves, J, prior.z, lambda)
-
-# Trim burn-in
-burnin <- n.sim.prior * burnin.prop
-prior.sample <- prior.chain$sample[-c(1:burnin)]
-
-# Prior Probs of Cluster Membership for each Area
-param.prior.zone <- 
[…]

### 2015-10-02 R/besag_newell.R

@@ -1,76 +1,0 @@
-besag_newell <-
-function(geo, population, cases, expected.cases=NULL, k, alpha.level){
-
-#-------------------------------------------------------------------------------
-# Initialization 
-#-------------------------------------------------------------------------------
-# If no expected.cases provided, set them if there are no expected counts
-if(is.null(expected.cases)){
-	p <- sum(cases)/sum(population)
-	expected.cases <- population*p
-}
-
-# geographical information computation
-geo.results <- zones(geo, population, 1)
-nearest.neighbors <- geo.results$nearest.neighbors
-distance <- geo.results$dist
-n.zones <- length(unlist(nearest.neighbors))
-
-
-#-------------------------------------------------------------------------------
-# Observed statistic computation
-#-------------------------------------------------------------------------------
-results <- besag_newell_internal(cases, expected.cases, nearest.neighbors, 
-                                 n.zones, k)
-
-# observed p.values for each areas
-p.values <- results$observed.p.values	
-# observed number of neighbors needed to observe k cases
-m.values <- results$observed.m.values	
-# actual observed number of cases
-k.values <- results$observed.k.values	
-
-# pick out areas that were significant and order them by p-value
-signif.indices <- order(p.values)[1:sum(p.values <= alpha.level)]
-
-# order remaining values
-signif.p.values <- p.values[signif.indices]
-signif.m.values <- m.values[signif.indices]
-signif.k.values <- k.values[signif.indices]
-
-
-# Create object to output
-# If none are significant, return NULL
-if(length(signif.indices) == 0){
-	clusters <- NULL
-} else {
-	clusters <- vector("list", length=length(signif.indices))
-
-	for( i in 1:length(clusters) ){	
-		# find areas included in cluster
-		cluster <- order(distance[signif.indices[i],])[1:signif.m.values[i]]
-		
-		new.cluster <- list(
-			location.IDs.included = cluster,
-			population = sum(population[cluster]),
-			number.of.cases = sum(cases[cluster]),
-			expected.cases = sum(expected.cases[cluster]),
-			SMR = sum(cases[cluster])/sum(expected.cases[cluster]),
-			p.value = signif.p.values[i]	
-		)
-		clusters[[i]] <- new.cluster
-	}
-}
-
-
-#-------------------------------------------------------------------------------
-# Output results
-#-------------------------------------------------------------------------------
-results <- list(
-	clusters=clusters,
-	p.values=p.values,
-	m.values=m.values,
-	observed.k.values=k.values
-)	
-return(results)
-}

### 2015-10-02 R/circle.R

@@ -1,37 +1,0 @@
-#' Compute cartesian coordinates of a cluster center and radius
-#' 
-#' This function is used for plotting purposes
-#' 
-#' @param geo A \code{n x 2} table of the x-coordinate and y-coordinates of the 
-#'   centroids of each area
-#' @param cluster.center The area index (an integer between \code{1} and
-#'   \code{n}) indicating the center of the circle
-#' @param cluster.end The area index (an integer between \code{1} and \code{n})
-#'   indicating the area at the end of the circle
-#'   
-#' @return cluster.radius A data frame that you can plot
-#'   
-#' @author Albert Y. Kim
-#' @seealso \code{\link{zones}}
-#'   
-#' @examples
-#' data(pennLC)
-#' geo <- pennLC$geo[,2:3]
-#' plot(geo,type='n')
-#' text(geo,labels=1:nrow(geo))
-#' lines( circle(geo, 23, 46), col = "red" ) 
-#'   
-circle <- function(geo, cluster.center, cluster.end) {
-  # Compute interpoint distance
-  distance <- dist(as.matrix(geo), upper = TRUE, diag = TRUE)
-  distance <- as.matrix(distance)[cluster.center, cluster.end]
-  
-  # For drawing radius of cluster on map
-  polar <- seq(0, 2 * pi, length = 1000)
-  
-  cluster.radius <- data.frame(cbind(x = distance * cos(polar) + geo$x[cluster.center], 
-    y = distance * sin(polar) + geo$y[cluster.center]))
-  
-  return(cluster.radius)
-}
-

### 2015-10-02 R/create_geo_objects.R

@@ -1,62 +1,0 @@
-create_geo_objects <-
-function(max.prop, population, centroids, sp.obj, area=NULL){
-
-# Number of areas
-n <- nrow(centroids)
-
-
-#-----------------------------------------------------------------------------
-# Set up single zones
-#-----------------------------------------------------------------------------
-if(max.prop < max(normalize(population))){
-  print(paste("max.prop needs to be at least", max(normalize(population))))
-}
-
-# Get basic single zone info
-geoInfo <- zones(centroids, population, max.prop)
-nearest.neighbors <- geoInfo$nearest.neighbors
-cluster.coords <- geoInfo$cluster.coords
-n.zones <- nrow(cluster.coords)
-
-# Create list of length n.zones indicating the component areas for each zone
-cluster.list <- vector(mode="list", length=n.zones)
-counter <- 1
-for(i in 1:n) {
-  nn <- nearest.neighbors[[i]]
-  for(j in 1:length(nn)) {
-    cluster.list[[counter]] <- nn[1:j] 
-    counter <- counter + 1  
-  } 
-}
-
-# If not provided, compute geographic surface area for each single zone.  
-if(is.null(area)){
-  area <- rep(0,n)
-  sp.obj.grid <- latlong2grid(sp.obj)
-  for(i in 1:n)
-    area[i] <- sp.obj.grid@polygons[[i]]@area
-  rm(sp.obj.grid)
-}
-areaz <- sapply(cluster.list,function(x){sum(area[x])})
-
-
-#-----------------------------------------------------------------------------
-# Generate overlap object which tracks the overlap between single zones
-#-----------------------------------------------------------------------------
-# For each area, get "presense":  list all unit zones that it is included in
-presence <- vector(mode="list",length=n)
-for(i in 1:n){
-  presence[[i]] <- which(sapply(cluster.list, function(x){is.element(i, x)}))
-}
-
-
-#----------------------------------------------------------------
-# Return list
-#----------------------------------------------------------------
-overlap <- list(presence=presence, cluster.list=cluster.list)
-return(list(
-  overlap=overlap, 
-  cluster.coords=cluster.coords, 
-  areaz=areaz)
-)
-}

### 2015-10-02 R/data.R

@@ -1,0 +1,14 @@
+#' Leukemia in Upstate New York
+#'
+#' Get data
+#'
+#' \itemize{
+#'   \item price. price in US dollars (\$326--\$18,823)
+#'   \item carat. weight of the diamond (0.2--5.01)
+#' }
+#'
+#' @docType data
+#' @keywords datasets
+#' @usage data(NYleukemia)
+#' @format A SpatialPolygonsDataFrame object
+"NYleukemia"

### 2015-10-02 R/eBayes.R

@@ -1,20 +1,0 @@
-eBayes <- function(Y, E, Xmat=NULL){
-	# Check for covariates
-	if (is.null(Xmat)) {
-		mod <- glm.nb(Y ~ 1+offset(log(E)), link=log)
-	}else{
-		mod <- glm.nb(Y ~ Xmat+offset(log(E)), link=log)
-	}
-
-	# Get MLE's of parameters
-	alpha <- mod$theta
-	muhat <- mod$fitted/E
-	
-	# Generate weights between global and local SMR
-	wgt <- E*muhat/(alpha+E*muhat); SMR <- Y/E
-	RR <- as.numeric(wgt*SMR + (1-wgt)*muhat)
-	RRmed <- qgamma(0.5,alpha+Y,(alpha+E*muhat)/muhat)
-	
-	# Output results
-	list(RR=RR, RRmed=RRmed, beta=mod$coeff, alpha=alpha, SMR=SMR)
-}

### 2015-10-02 R/estimate_lambda.R

@@ -1,50 +1,0 @@
-estimate_lambda <-
-function(n.sim, J, prior.z, overlap, pi0, equal.non.null=FALSE){
-
-n.zones <- length(prior.z)
-
-# q_0 = q_1 = 1, since there is no concept of overlap for j=0, 1
-q <- c(1, 1, rep(0, J-1))
-var.q <- c(NA, NA, rep(0, J-1))
-
-
-#-------------------------------------------------------------------------------
-# All k's, sample configurations of k single zones, and see if the k single
-# zones overlap or not
-#-------------------------------------------------------------------------------
-for(k in 2:J){
-  unit.zones <- sample(1:n.zones, k*n.sim, replace=TRUE, prob=prior.z)
-  unit.zones <- matrix(unit.zones, ncol=k)
-  no.overlap <- check_overlap(unit.zones, overlap)
-  
-  denom <- rep(factorial(k), n.sim)
-  
-  # Compute multinomial coefficient watching out for when certain single zones
-  # are sampled more than once
-  duplicates <- apply(unit.zones, 1, function(x){length(unique(x))})
-  duplicates <- which(duplicates != k)
-  for(i in duplicates) {
-    denom[i] <-  factorial(k)/prod(factorial(table(unit.zones[i,])))
-  }
-  
-  q[k+1] <- mean(no.overlap/denom)
-  var.q[k+1] <- var(no.overlap/denom)/n.sim
-}
-
-
-#-------------------------------------------------------------------------------
-# Using q compute estimate of lambda and prior probability of j clusters
-#-------------------------------------------------------------------------------
-if(equal.non.null) {
-  lambda <- (1-pi0)/q
-  lambda[1] <- pi0*J  # null prior
-  lambda <- normalize(lambda)
-} else {
-  lambda <- (1-pi0)/((1-pi0)*J + pi0*sum(q[-1]))
-  lambda <- rep(lambda, J)
-  lambda <- c(1-sum(lambda), lambda)
-}
-prior.j <- normalize(lambda*q)
-
-return(list(lambda=lambda, prior.j=prior.j, q=q))
-}

### 2015-10-02 R/expected.R

@@ -1,27 +1,0 @@
-expected <-
-function(population, cases, n.strata){
-
-n <- length(population)/n.strata
-E <- rep(0, n)
-qNum <- rep(0, n.strata)
-qDenom <- rep(0, n.strata)
-q <- rep(0, n.strata)
-
-
-# Compute q: strata-specific rates. Numerator and denominator separately
-for(i in 1:n.strata){
-  indices <- rep(i, n) + seq(0, n-1)*n.strata
-  qNum[i] <- qNum[i] + sum(cases[indices])
-  qDenom[i] <- qDenom[i] + sum(population[indices])
-}
-q <- qNum/qDenom
-
-
-# Compute E expected counts
-for(i in 1:n) {
-  indices <- 1:n.strata + (i-1)*n.strata
-  E[i] <- sum(population[indices]*q)
-}
-
-return(E)
-}

### 2015-10-02 R/grid2latlong.R

@@ -1,48 +1,0 @@
-grid2latlong <-
-function(input){
-
-toradians <- atan(1)/45
-radiusearth <- 0.5*(6378.2+6356.7)
-sine51 <- sin( 51.5*toradians )
-
-
-#-------------------------------------------------------------------------------
-# If a Spatial Polygons
-#-------------------------------------------------------------------------------
-if(is(input)[1] == "SpatialPolygons"){
-  for(i in 1:length(input@polygons)){
-    # for all Polygons's in polygon
-    for(j in 1:length(input@polygons[[i]]@Polygons)){
-      # Convert coordinates
-      new.coords <- as.matrix(cbind(
-        input@polygons[[i]]@Polygons[[j]]@coords[,1]/(toradians*radiusearth*sine51),
-        input@polygons[[i]]@Polygons[[j]]@coords[,2]/(toradians*radiusearth)
-      ))
-      colnames(new.coords) <- NULL
-      rownames(new.coords) <- NULL
-      
-      # Update Polygons
-      input@polygons[[i]]@Polygons[[j]]@coords <- new.coords
-      input@polygons[[i]]@Polygons[[j]] <- Polygon(input@polygons[[i]]@Polygons[[j]])
-    }
-    # Update polygons
-    input@polygons[[i]] <- Polygons(
-      input@polygons[[i]]@Polygons,
-      ID=input@polygons[[i]]@ID
-    )	
-  }
-  output <- SpatialPolygons(input@polygons,proj4string=CRS("+proj=utm"))
-
-  
-#-------------------------------------------------------------------------------
-# else return numeric
-#-------------------------------------------------------------------------------
-}else{
-  output <- data.frame(cbind(
-    x=input[, 1]/(toradians*radiusearth*sine51),
-    y=input[, 2]/(toradians*radiusearth)
-  ))		
-}
-
-return(output)	
-}

### 2015-10-02 R/latlong2grid.R

@@ -1,50 +1,0 @@
-latlong2grid <-
-function(input){
-  
-toradians <- atan(1)/45
-radiusearth <- 0.5*(6378.2+6356.7)
-sine51 <- sin( 51.5*toradians )
-
-
-#-------------------------------------------------------------------------------
-# If a Spatial Polygon
-#-------------------------------------------------------------------------------
-if(any(is(input) == "SpatialPolygons")){ 
-  for( i in 1:length(input@polygons) ){
-    # for all Polygons's in polygon
-    for( j in 1:length(input@polygons[[i]]@Polygons) ){
-      # Convert coordinates
-      new.coords <- cbind(
-        (input@polygons[[i]]@Polygons[[j]]@coords[,1]*toradians)*radiusearth*sine51,
-        (input@polygons[[i]]@Polygons[[j]]@coords[,2]*toradians)*radiusearth
-      )
-      new.coords <- as.matrix(new.coords)
-      colnames(new.coords) <- NULL
-      rownames(new.coords) <- NULL
-      
-      # Update Polygons
-      input@polygons[[i]]@Polygons[[j]]@coords <- new.coords
-      input@polygons[[i]]@Polygons[[j]] <- Polygon(input@polygons[[i]]@Polygons[[j]])
-    }
-    # Update polygons
-    input@polygons[[i]] <- Polygons(
-      input@polygons[[i]]@Polygons,
-      ID=input@polygons[[i]]@ID
-    )	
-  }
-  
-  output <- SpatialPolygons(input@polygons,proj4string=CRS("+proj=utm"))
-
-  
-#-------------------------------------------------------------------------------  
-# else return numeric
-#-------------------------------------------------------------------------------
-}else{
-  output <- data.frame(cbind(
-    x=(input[,1]*toradians)*radiusearth*sine51,
-    y=(input[,2]*toradians)*radiusearth
-  ))		
-}
-
-return(output)	
-}

### 2015-10-02 R/plot_crayons.R

@@ -1,65 +1,0 @@
-#' Illustration of crayon colors
-#'
-#' Creates a plot of the crayon colors in \code{\link{brocolors}}
-#'
-#' @param method2order method to order colors (\code{"hsv"} or \code{"cluster"})
-#' @param cex character expansion for the text
-#' @param mar margin paramaters; vector of length 4 (see \code{\link[graphics]{par}})
-#'
-#' @return None
-#'
-#' @author Karl W Broman, \email{kbroman@@biostat.wisc.edu}
-#' @references \url{http://en.wikipedia.org/wiki/List_of_Crayola_crayon_colors}
-#' @seealso \code{\link{brocolors}}
-#' @keywords hplot
-#'
-#' @examples
-#' plot_crayons()
-#'
-#' @export
-#' @importFrom grDevices rgb2hsv
-#' @importFrom graphics par plot rect text
-#'
-plot_crayons <-
-  function(method2order=c("hsv", "cluster"), cex=0.6, mar=rep(0.1, 4))
-  {
-    method2order <- match.arg(method2order)
-    
-    crayons <- brocolors("crayons")
-    
-    # get rgb
-    colval <- col2rgb(crayons)
-    
-    if(method2order == "hsv") {
-      # convert to hsv
-      colval <- t(rgb2hsv(colval))
-      
-      # order the colors; first two lines are to get black/gray/silver/white first
-      ord <- order(names(crayons)!="Black", names(crayons)!="Gray",
-        names(crayons)!="Silver", names(crayons)!="White",
-        colval[,1], colval[,2], colval[,3])
-      
-    } else {
-      ord <- hclust(dist(t(colval)))$ord
-    }
-    
-    oldmar <- par("mar")
-    on.exit(par(mar=oldmar))
-    
-    par(mar=mar)
-    x <- (1:7)-1
-    y <- (1:19)-1
-    x <- rep(x, each=19)
-    y <- rep(y, 7)
-    
-    plot(0, 0, type="n", xlab="", ylab="", xaxs="i", yaxs="i",
-      xlim=c(0, max(x)+1), ylim=c(max(y)+0.5, -0.5),
-      xaxt="n", yaxt="n")
-    
-    dx <- 0.2
-    dy <- 0.4
-    rect(x+dx/4, y-dy, x+dx, y+dy, border="black",
-      col=crayons[ord])
-    
-    text(x+dx*1.2, y, names(crayons)[ord], cex=cex, adj=c(0, 0.5))
-  }

### 2015-10-02 R/plotmap.R

@@ -1,44 +1,0 @@
-plotmap <-
-function(values, map, log=FALSE, nclr=7, include.legend=TRUE, lwd=0.5, round=3,
-         brks=NULL, legend=NULL, location='topright', rev=FALSE, leg.cex=1){
-
-# create colors, each based on quantiles of data
-plotclr <- grey(1-seq(0, 1, by=1/(nclr-1)))
-
-nclr <- nclr + 1
-if(is.null(brks)){
-  if(log){
-    brks <- exp(
-      seq(from=min(log(values)), to=max(log(values)), length.out=nclr)
-      )
-  }else{
-    brks <- seq(from=min(values), to=max(values), length.out=nclr)
-  }
-}
-nclr <- nclr - 1
-
-
-# Assign colors based on intervals
-if(rev){
-  plotclr <- rev(plotclr)
-}
-colornum <- findInterval(values, brks, all.inside=T);
-colcode <- plotclr[colornum]
-
-
-# Modify legend
-if(is.null(legend)){
-  legend <- paste(
-    signif(brks, digits=round)[1:nclr],
-    signif(brks, digits=round)[2:(nclr+1)],
-    sep=" - "
-  )
-}
-
-
-plot(map, axes=TRUE, col=colcode, lwd=lwd)
-if(include.legend) {
-  legend(location, legend = legend, fill = plotclr, bty = "n", cex=leg.cex, 
-         pt.cex = 1)}
-
-}

### 2015-10-02 R/polygon2spatial_polygon.R

@@ -1,72 +1,0 @@
-polygon2spatial_polygon <-
-function(poly, coordinate.system, area.names = NULL, nrepeats = NULL){
-
-#-------------------------------------------------------------------------------
-# Deal with non-specified values
-#-------------------------------------------------------------------------------
-if(missing(coordinate.system)){
-	stop("Coordinate system must be specified: '+proj=utm' or '+proj=longlat'.")
-}
-
-if(is.null(nrepeats)){
-	nrepeats <- rep(1, sum(is.na(poly[,1]))+1 )	
-}
-
-if(is.null(area.names)){
-	area.names <- as.character( 1:length(nrepeats) )
-}
-
-
-#-------------------------------------------------------------------------------
-# Create list of all polygon objects
-#-------------------------------------------------------------------------------
-na.index <- which(is.na(poly[,1]))
-n <- length(nrepeats)
-list.polygon <- NULL
-
-# First Case
-list.polygon <-	list(Polygon(poly[1:(na.index[1]-1),], hole=FALSE))					
-
-# Middle cases
-for(i in 1:(length(na.index)-1)){
-	list.polygon <- c(list.polygon,list(Polygon(
-		poly[(na.index[i]+1):(na.index[i+1]-1),], hole=FALSE)))	
-}
-
-# Last case
-list.polygon <-	c(list.polygon,list(Polygon(
-  poly[(na.index[i+1]+1):length(poly[,1]),], hole=FALSE)
-  ))
-
-
-#-------------------------------------------------------------------------------
-# From list of polygon objects, create "polygon" objects, that has one element
-# for each county.  A county can consist of several polygon as indicated by
-# nrepeats
-#-------------------------------------------------------------------------------
-list.polygons <- NULL
-
-start <- 1
-for( i in 1:length(nrepeats) ){
-	end <- start + nrepeats[i] - 1
-	
-	temp.polygon <- NULL
-	for(j in start:end){
-		temp.polygon <- c(temp.polygon, list(list.polygon[[j]]))
-	}
-	
-	list.polygons <- c(list.polygons, list(
-		Polygons(temp.polygon, ID=area.names[i])
-	))
-	start <- end + 1	
-}
-
-
-#-------------------------------------------------------------------------------
-# Output spatial polygons object
-#-------------------------------------------------------------------------------
-Spatial.Polygon <- 
-  SpatialPolygons(list.polygons, proj4string=CRS(coordinate.system))
-
-return(Spatial.Polygon)
-}

### 2015-10-02 R/process_MCMC_sample.R

@@ -1,61 +1,0 @@
-process_MCMC_sample <-
-function(sample, param, RR.area, cluster.list, cutoffs){
-
-n <- length(RR.area)
-n.sim <- length(sample)
-  
-# Store outputs here. These are averages weighted by the probability of 
-# each configuration
-high.area <- low.area <- RR.est.area <- rep(0, n)
-
-# Zone based measures
-high.z <- pgamma(cutoffs$high, param$shape, param$rate, lower.tail=FALSE)
-low.z <- pgamma(cutoffs$low, param$shape, param$rate, lower.tail=TRUE)
-RR.z <- param$shape / param$rate
-
-
-#-------------------------------------------------------------------------------
-# Find probabilities of cluster and anti-cluster membership & estimated 
-# relative risk
-#-------------------------------------------------------------------------------
-# Get probabilities of each configuration
-unique <- unique(sample)
-indices <- match(sample, unique)
-prob <- table(indices)/n.sim
-prob.names <- as.numeric(names(prob))
-
-# Over all sampled configurations
-for(i in 1:length(prob.names)){
-  config <- unique[[prob.names[i]]]
-  k <- length(config)
-  
-  if(k==0){
-    RR.est.area <- RR.est.area + RR.area * prob[i]
-  }else{
-    # Used to identify all areas included in configuration
-    config.areas <- NULL    
-    for(j in 1:k){        
-      # All areas in this single zone in this configuration
-      zone <- config[j]
-      areas <- cluster.list[[zone]]
-      config.areas <- c(config.areas, areas)
-      
-      # Update all areas inside configuration
-      high.area[areas] <- high.area[areas] + high.z[zone] * prob[i]
-      low.area[areas] <- low.area[areas] + low.z[zone] * prob[i]
-      RR.est.area[areas] <- RR.est.area[areas] + RR.z[zone] * prob[i] 
-    }
-    
-    # Update all areas outside configuration
-    outside.areas <- setdiff(1:n, config.areas)
-    RR.est.area[outside.areas] <- 
-      RR.est.area[outside.areas] + RR.area[outside.areas]*prob[i]
-  }
-}
-
-return(list(
-  high.area=high.area,
-  low.area=low.area,
-  RR.est.area=RR.est.area)
-  )
-}

### 2015-10-02 R/zones.R

@@ -1,55 +1,83 @@
-zones <-
-function(geo, population, pop.upper.bound){
+# When using dplyr, R CMD CHECK returns NOTE that there is "no visible binding 
+# for global variable", hence this hack fix.
+# no visible binding for global variable
+utils::globalVariables(c("distance", "prop_pop", "cum_prop", "."))
 
-# number of areas
-n <- nrow(geo)
-# total population
-total.pop <- sum(population)
-# Interpoint distance matrix
-dist <- as.matrix(dist(as.matrix(geo), upper=TRUE, diag=TRUE))
-
-
-#-------------------------------------------------------------------------------
-# For each area, list of closest neighbors up until pop.upper.bound of
-# population is met.  We count the number of candidate zones in n.zones 
-# Note:  for each county, the list of nearest neighbors form the total number of
-# candidate zones
-#-------------------------------------------------------------------------------
-nearest.neighbors <- vector(mode="list", length=n)
-
-n.zones <- 0
-vector.cutoffs <- rep(0, n+1)
-
-for(i in 1:n) {	
-  # Sort the areas by distance, then include them one-by-one until cluster bound
-  # is met
-  neighbors <- order(dist[,i])
+#' Define a study region's single zones.
+#' 
+#' Using the centroids of the n areas in a study region, define the set of 
+#' single zones based on the upper bound \code{pop_upper_bound} of the 
+#' proportion of the study region's population each single zone can contain.
+#' 
+#' @param centroids A data frame with 2 columns of the centroids of the n areas 
+#'   in the study region.
+#' @param population A vector of length n of the corresponding population 
+#'   counts.
+#' @param pop_upper_bound The upper bound of the proportion of the study 
+#'   region's population the single zones can contain.
+#'   
+#' @return A list with two objects \describe{ 
+#'   \item{\code{nearest_neighbors}}{A list of length n of the single zones 
+#'   for each area} 
+#'   \item{\code{cluster_coords}}{A data frame with 2 columns of the centering 
+#'   and radial area for each single zone.} 
+#'   }
+#' @references Kulldorff, M. (1997) A spatial scan statistic. 
+#'   \emph{Communications in Statistics: Theory and Methods}, \bold{26}, 
+#'   1481--1496.
+#' @export
+#'   
+#' @examples 
+#' data(NYleukemia)
[…]

### 2015-10-02 src/RcppExports.cpp

@@ -1,236 +1,0 @@
-// This file was generated by Rcpp::compileAttributes
-// Generator token: 10BE3573-1514-4C36-9D1C-5A225CD40393
-
-#include <RcppArmadillo.h>
-#include <Rcpp.h>
-
-using namespace Rcpp;
-
-// normalize
-NumericVector normalize(NumericVector x);
-RcppExport SEXP SpatialEpi_normalize(SEXP xSEXP) {
-BEGIN_RCPP
-    Rcpp::RObject __result;
-    Rcpp::RNGScope __rngScope;
-    Rcpp::traits::input_parameter< NumericVector >::type x(xSEXP);
-    __result = Rcpp::wrap(normalize(x));
-    return __result;
-END_RCPP
-}
-// NumericVectorEquality
-int NumericVectorEquality(NumericVector x, NumericVector y);
-RcppExport SEXP SpatialEpi_NumericVectorEquality(SEXP xSEXP, SEXP ySEXP) {
-BEGIN_RCPP
-    Rcpp::RObject __result;
-    Rcpp::RNGScope __rngScope;
-    Rcpp::traits::input_parameter< NumericVector >::type x(xSEXP);
-    Rcpp::traits::input_parameter< NumericVector >::type y(ySEXP);
-    __result = Rcpp::wrap(NumericVectorEquality(x, y));
-    return __result;
-END_RCPP
-}
-// ldnbinom
-double ldnbinom(double y, double E, double a, double b);
-RcppExport SEXP SpatialEpi_ldnbinom(SEXP ySEXP, SEXP ESEXP, SEXP aSEXP, SEXP bSEXP) {
-BEGIN_RCPP
-    Rcpp::RObject __result;
-    Rcpp::RNGScope __rngScope;
-    Rcpp::traits::input_parameter< double >::type y(ySEXP);
-    Rcpp::traits::input_parameter< double >::type E(ESEXP);
-    Rcpp::traits::input_parameter< double >::type a(aSEXP);
-    Rcpp::traits::input_parameter< double >::type b(bSEXP);
-    __result = Rcpp::wrap(ldnbinom(y, E, a, b));
-    return __result;
-END_RCPP
-}
-// ldmultinom
-double ldmultinom(NumericVector x, NumericVector prob);
-RcppExport SEXP SpatialEpi_ldmultinom(SEXP xSEXP, SEXP probSEXP) {
-BEGIN_RCPP
-    Rcpp::RObject __result;
-    Rcpp::RNGScope __rngScope;
[…]

### 2015-10-02 src/bayes_cluster.cpp

@@ -1,762 +1,0 @@
-#include <RcppArmadilloExtensions/sample.h>
-// [[Rcpp::depends(RcppArmadillo)]]
-using namespace Rcpp;
-
-
-
-// [[Rcpp::export]]
-NumericVector normalize(NumericVector x) {
-  int n = x.size();
-  NumericVector y(n);
-  double sum_x = sum(x);
-  for(int i=0; i<n; ++i){
-    y[i] = x[i]/sum_x;
-	}
-  return y;
-}
-
-
-
-// [[Rcpp::export]]
-int NumericVectorEquality(NumericVector x, NumericVector y){  
-  if(x.size() != y.size()) {
-    return 0;
-  }
-
-  int vector_equal = 1;
-  for(int i=0; i<x.size(); ++i) {
-    if(x[i] != y[i]) {
-      vector_equal = 0;
-      break;
-    }
-  }
- 
- return vector_equal;
-}
-
-
-
-// [[Rcpp::export]]
-double ldnbinom(double y, double E, double a, double b) {
-  double value = lgamma(y+a) - lgamma(y+1) - lgamma(a) + y*(log(E)-log(E+b)) +
-  a*(log(b)-log(E+b));
-
-  return value;
-}
-
-
-
-// [[Rcpp::export]]
-double ldmultinom(NumericVector x, NumericVector prob) {
-  int len = x.size();
-	NumericVector p(len);
-	double value = 0, sum_p = sum(prob);
-
-	// normalize probabilities
-	for(int i=0; i<len; ++i) {
-		p[i] = prob[i]/sum_p;
-	}
-
-	value = lgamma(sum(x) + 1);
-	for(int i=0; i<len; ++i) {
-		if(p[i] != 0) {
-			value -= lgamma(x[i] + 1);
-			value += x[i] * log(p[i]);
-		}
-	}
-
-	return value;
-}
-
-
-
-// [[Rcpp::export]]
-NumericVector coeff(NumericVector y_vector, NumericVector E_vector, 
-	NumericVector a_values, NumericVector b_values, List cluster_list) {
-	
-	double yz, Ez, sum_log_fy_inside;
-	int n = y_vector.size(), n_zones = cluster_list.size(), index;
-	NumericVector log_fy_n(n), coeff(n_zones), y_temp(n), E_temp(n), inside(n);
-
-	// Compute n normalization terms (log negbinom)
-	for(int i=0; i<n; ++i) {
[…]

### 2015-10-02 src/cluster_detection.cpp

@@ -1,155 +1,0 @@
-#include <Rcpp.h>
-using namespace Rcpp;
-
-
-// [[Rcpp::export]]
-double binomialLogLkhd(double cz, double nz, double N, double C) {
-	double logLkhd = 0;
-
-	if(cz / nz <= (C - cz)/(N - nz)) {
-		logLkhd = 0;
-	} else {
-		logLkhd  =	
-		N * ( log( N-nz-C+cz ) - log( N-C )  + log( N ) - log( N-nz ) ) +
-		nz * ( log( nz-cz ) - log( N-nz-C+cz )  + log( N-nz ) - log( nz ) )+
-		cz * ( log( cz ) - log( nz-cz ) + log( N-nz-C+cz ) - log( C-cz ) )+
-		C * ( log( C-cz ) - log( C ) + log( N-C ) - log( N-nz-C+cz ) );
-	}
-
-	return logLkhd;
-}
-
-
-// [[Rcpp::export]]
-double poissonLogLkhd(double cz, double nz, double N, double C) {
-	double logLkhd = 0;
-
-	if(cz / nz <= (C - cz)/(N - nz)) {
-		logLkhd = 0;
-	} else {
-		logLkhd =
-		cz * log(  (cz / nz) )  +
-		cz * log(  ( (N - nz)/( C - cz) ) )  +
-		C * log(  ( (C-cz)/(N-nz) )  ) +
-		C * log(  ( N/ C )  );
-	}  
-	return logLkhd;
-}
-
-
-// [[Rcpp::export]]
-NumericVector computeAllLogLkhd(
-	NumericVector observedCases, NumericVector expectedCases, 
-	List nearestNeighborsList, int nZones, String logLkhdType) {
-
-	NumericVector allLogLkhd(nZones);
-	int nAreas = expectedCases.size(), C = sum(observedCases), 
-	N = sum(expectedCases), index = 0;
-	int nNeighbors = 0;
-	double cz = 0.0, nz = 0.0;
-
-	for (int i = 0; i < nAreas; ++i) {
-		cz = 0;
-		nz = 0;
-
-		Rcpp::NumericVector nearestNeighbors = nearestNeighborsList[i];
-		nNeighbors =  nearestNeighbors.size();
-
-		// For each area's nearest neighbors
-		for(int j = 0; j < nNeighbors; ++j) { 
-			// Watch off by 1 vector indexing in C as opposed to R
[…]

### 2015-10-02 testing.R

@@ -1,0 +1,25 @@
+# Testing define_zones
+load("/Users/rudeboybert/Dropbox/SpatialEpi/data/NYleukemia.rda")
+geo <- NYleukemia$geo[,2:3] %>% as.data.frame()
+population <- NYleukemia$data$population
+pop_upper_bound <- 0.15
+
+testing <- define_zones(geo, population, pop_upper_bound)
+testing2 <- zones(geo, population, pop_upper_bound)
+
+
+
+# Save NYleukemia data as SpatialPolygonsDataFrame
+load("/Users/rudeboybert/Dropbox/SpatialEpi/data/NYleukemia.rda")
+
+sp <- NYleukemia$spatial.polygon
+data <- NYleukemia$data %>% 
+  rename(FIPS = censustract.FIPS) %>% 
+  mutate(ID = 1:nrow(NYleukemia$data)) %>% 
+  select(ID, FIPS, population, cases)
+
+NYleukemia <- SpatialPolygonsDataFrame(sp, data)
+devtools::use_data(NYleukemia, NYleukemia, overwrite=TRUE)
+
+library(sp)
+testing <- define_zones(coordinates(NYleukemia), NYleukemia$population, 0.15)

### 2015-10-16 R/kulldorff.R
* Added instructions for installing rgeos and rdgal * Started mapping variables

@@ -11,10 +11,11 @@
 #' @param all_log_lkhd 
 #' @param lambdas 
 #'
-#' @return
+#' @return Cluster report
 #' @export
 #'
 #' @examples
+#' 1+1
 get_cluster_report <- function(cluster_index, population, cases, expected_cases,
                                zone_info, all_log_lkhd, lambdas){
   cluster <- return_single_zone_areas(cluster_index, zone_info)

### 2015-10-16 vignettes/SpatialEpi.R

@@ -1,3 +1,37 @@
+## ---- eval=FALSE---------------------------------------------------------
+#  library(rgdal)
+#  library(rgeos)
+
+## ---- message=FALSE, collapse=TRUE, tidy=TRUE,tidy.opts=list(width.cutoff=40)----
+library(maptools)
+library(dplyr)
+library(spdep)
+# Awkward
+data(nc.sids)
+sids <- data.frame(
+  Observed=nc.sids$SID74,
+  Expected=nc.sids$BIR74*sum(nc.sids$SID74)/sum(nc.sids$BIR74),
+  x=nc.sids$x, 
+  y=nc.sids$y)
+
+nc.sids <- readShapePoly(
+  fn = system.file("etc/shapes/sids.shp", package="spdep")[1],
+  ID="FIPSNO",
+  proj4string=CRS("+proj=longlat +ellps=clrk66"))
+
+# Create a copy to modify
+nc.sids@data <- nc.sids@data %>% 
+  mutate(
+    Y = SID74,
+    E = sum(Y) * BIR74/sum(BIR74),
+    SMR74 = Y/E,
+    EXP74 = E
+  )
+brks <- seq(0, 5)
+
+## ---- echo=TRUE, collapse=TRUE,fig.height=4.5,fig.width=4, fig.cap="Map of SMRs for SIDS in 1974 in North Carolina", tidy.opts=list(width.cutoff=35)----
+spplot(nc.sids, "SMR74", at=brks, col.regions=grey.colors(5,start=.9,end=.1))
+
 ## ---- fig.show='hold'----------------------------------------------------
 plot(1:10)
 plot(10:1)

### 2015-10-23 R/RcppExports.R
Done first pass of map_variable function.

@@ -24,12 +24,11 @@ poissonLogLkhd <- function(cz, nz, C, N) {
 
 #' Compute log likelihood ratio test statistic for all single zones.
 #' 
-#' @param observedCases
-#' @param expectedCases
-#' @param nearestNeighborsList
-#' @param nZones
-#' @param logLkhdType
-#' @return Vector of log likelihood
+#' @param observedCases vector of observed cases
+#' @param expectedCases vector of expected cases
+#' @param nearestNeighborsList list of nearest neighbors
+#' @param logLkhdType type of log-likelihood 
+#' @return Vector of log likelihoods for each single zone
 #' @export
 computeAllLogLkhd <- function(observedCases, expectedCases, nearestNeighborsList, logLkhdType) {
     .Call('SpatialEpi_computeAllLogLkhd', PACKAGE = 'SpatialEpi', observedCases, expectedCases, nearestNeighborsList, logLkhdType)
@@ -38,7 +37,7 @@ computeAllLogLkhd <- function(observedCases, expectedCases, nearestNeighborsList
 #' Compute log likelihood ratio test statistic for all 
 #' 
 #' @inheritParams computeAllLogLkhd
-#' @param permutedCaseMatrix blah
+#' @param permutedCaseMatrix matrix of null simulated cases
 #' @return Max
 #' @export
 kulldorffMC <- function(permutedCaseMatrix, expectedCases, nearestNeighborsList, logLkhdType) {

### 2015-10-23 R/SpatialEpi.R

@@ -8,6 +8,8 @@
 #' @docType package
 #' @import dplyr
 #' @import sp
+#' @import maptools
+#' @import ggplot2
 #' @name SpatialEpi
 #' @useDynLib SpatialEpi
 #' @importFrom Rcpp sourceCpp

### 2015-10-23 R/geographic_objects.R

@@ -1,7 +1,8 @@
-# When using dplyr, R CMD CHECK returns NOTE that there is "no visible binding 
-# for global variable", hence this hack fix.
-# no visible binding for global variable
+# R CMD CHECK returns NOTE that there is "no visible binding 
+# for global variable" for variables used by dplyr/ggplot2, hence this hack fix.
 utils::globalVariables(c("distance", "prop_pop", "cum_prop", "."))
+
+
 
 #' Define a study region's single zones.
 #' 
@@ -9,11 +10,11 @@ utils::globalVariables(c("distance", "prop_pop", "cum_prop", "."))
 #' single zones based on the upper bound \code{pop_upper_bound} of the 
 #' proportion of the study region's population each single zone can contain.
 #' 
-#' @param centroids A data frame with 2 columns of the centroids of the n areas 
+#' @param centroids data frame with 2 columns of the centroids of the n areas 
 #'   in the study region.
-#' @param population A vector of length n of the corresponding population 
+#' @param population vector of length n of the corresponding population 
 #'   counts.
-#' @param pop_upper_bound The upper bound of the proportion of the study 
+#' @param pop_upper_bound upper bound of the proportion of the study 
 #'   region's population the single zones can contain.
 #'   
 #' @return A list with two objects \describe{ 
@@ -25,6 +26,7 @@ utils::globalVariables(c("distance", "prop_pop", "cum_prop", "."))
 #' @references Kulldorff, M. (1997) A spatial scan statistic. 
 #'   \emph{Communications in Statistics: Theory and Methods}, \bold{26}, 
 #'   1481--1496.
+#' @seealso \code{\link{create_geo_objects}}
 #' @export
 #'   
 #' @examples 
@@ -78,20 +80,22 @@ define_single_zones <- function(centroids, population, pop_upper_bound) {
 }
 
 
+
 #' Obtain all areas included in a single zone
 #'
-#' @param cluster_index integer index of single zone
-#' @param zone_info output of \code{\link{define_single_zones}} function
+#' @param single_zone_index integer index of single zone
+#' @param zone_info output of \code{\link{define_single_zones}} defining single
+#'   zones
 #'
 #' @return vector of areas
 #' @export
 #'
 #' @examples
 #' 1+1
-return_single_zone_areas <- function(cluster_index, zone_info){
+return_single_zone_areas <- function(single_zone_index, zone_info){
   # Obtain single zone center and radial areas
[…]

### 2015-10-23 R/kulldorff.R

@@ -1,46 +1,81 @@
 #' Generate cluster reports
 #'
-#' Return a report on most likely cluster and secondary clusters containing the
-#' same fields as SatScan
+#' Return a report for a given cluster containing the same information as SatScan
 #'
-#' @param cluster_index 
-#' @param population 
-#' @param cases 
-#' @param expected_cases 
-#' @param zone_info 
-#' @param all_log_lkhd 
-#' @param lambdas 
+#' @inheritParams define_single_zones
+#' @inheritParams return_single_zone_areas
+#' @inheritParams kulldorff
+#' @param all_log_lkhd vector of the log likelihood of each single zone
+#' @param lambdas vector of simulated max log likelihood
 #'
-#' @return Cluster report
+#' @return For a particular clusters a report including \describe{ 
+#'   \item{\code{location_IDs_included}}{ID's of areas in cluster, in order of distance} 
+#'   \item{\code{population}}{population of cluster} 
+#'   \item{\code{number_of_cases}}{number of cases in cluster}
+#'   \item{\code{expected_cases}}{expected number of cases in cluster}
+#'   \item{\code{SMR}}{estimated SMR of cluster}
+#'   \item{\code{log_lkhd_ratio}}{log-likelihood of cluster}
+#'   \item{\code{monte_carlo_rank}}{rank of lkhd of cluster within Monte Carlo simulated values}
+#'   \item{\code{p_value}}{Monte Carlo \eqn{p}-value}
+#'   }
+#' 
 #' @export
 #'
 #' @examples
 #' 1+1
-get_cluster_report <- function(cluster_index, population, cases, expected_cases,
-                               zone_info, all_log_lkhd, lambdas){
-  cluster <- return_single_zone_areas(cluster_index, zone_info)
+get_cluster_report <- function(single_zone_index, zone_info, cases, population,  
+                               expected_cases, all_log_lkhd, lambdas){
+  # Get areas included in single zone
+  cluster_areas <- return_single_zone_areas(single_zone_index, zone_info)
   
+  # Return information
   output <- list(
-    location_IDs_included = cluster,
-    population = sum(population[cluster]),	
-    number_of_cases = sum(cases[cluster]),
-    expected_cases = sum(expected_cases[cluster]),
-    SMR = sum(cases[cluster])/sum(expected_cases[cluster]),
-    log_lkhd_ratio = all_log_lkhd[cluster_index], 
[…]

### 2015-10-23 R/mapping.R

@@ -1,0 +1,66 @@
+# R CMD CHECK returns NOTE that there is "no visible binding 
+# for global variable" for variables used by dplyr/ggplot2, hence this hack fix.
+utils::globalVariables(c("lat", "long", "group", "."))
+
+
+#' Plot ggplot compliant map
+#' 
+#' As described in \url{https://github.com/hadley/ggplot2/wiki/plotting-polygon-shapefiles}
+#'
+#' @param sp_obj SpatialPolygonsDataFrame object
+#' @param variable Variable within data frame \code{sp_obj@@data}
+#' @param x_label Label for eastings
+#' @param y_label Label for northings
+#' @param title Plot title
+#' @param legend_label label for legend. Defaults to variable name.
+#'
+#' @return a ggplot object
+#' @export
+#'
+#' @examples
+#' shapefile_name <- system.file("extdata/sids/sids.shp", package="SpatialEpi")
+#' proj4string <- sp::CRS("+proj=longlat +ellps=clrk66")
+#' nc_sids <- maptools::readShapePoly(
+#'    fn = shapefile_name,
+#'    ID = "FIPSNO",
+#'    proj4string = proj4string
+#'    )
+#' map_variable(nc_sids)
+#' names(nc_sids)
+#' nc_sids$RATE74 <- nc_sids$SID74/nc_sids$BIR74
+#' nc_map <- map_variable(nc_sids, "RATE74")
+#' nc_map
+map_variable <- function(sp_obj, variable = NULL, x_label = "eastings", 
+                         y_label = "northings", title = NULL, 
+                         legend_label = NULL){
+  if(is.null(legend_label)){
+    legend_label <- variable
+  }
+  
+  # Create data frame from SpatialPolygonsDataFrame object for ggplot'ing
+  sp_obj@data$id <- rownames(sp_obj@data)
+  map_points <- fortify(sp_obj, region="id")
+  map_df <- inner_join(map_points, sp_obj@data, by="id")
+  
+  # Set up base plot
+  map_ggplot <- 
+    ggplot(map_df, aes(x=long, y=lat, group=group)) +
+    coord_map() +
+    theme_bw() +
+    xlab(x_label) + 
+    ylab(y_label) + 
+    ggtitle(title) + 
+    geom_polygon(fill="white") + 
+    geom_path(col="black", size=0.5)
+  
+  # If specified, plot variable
+  if(!is.null(variable)) {
+    map_ggplot <- map_ggplot + 
[…]

### 2015-10-23 src/kulldorff.cpp

@@ -49,12 +49,11 @@ double poissonLogLkhd(double cz, double nz, double C, double N) {
 
 //' Compute log likelihood ratio test statistic for all single zones.
 //' 
-//' @param observedCases
-//' @param expectedCases
-//' @param nearestNeighborsList
-//' @param nZones
-//' @param logLkhdType
-//' @return Vector of log likelihood
+//' @param observedCases vector of observed cases
+//' @param expectedCases vector of expected cases
+//' @param nearestNeighborsList list of nearest neighbors
+//' @param logLkhdType type of log-likelihood 
+//' @return Vector of log likelihoods for each single zone
 //' @export
 // [[Rcpp::export]]
 NumericVector computeAllLogLkhd(NumericVector observedCases, 
@@ -107,7 +106,7 @@ NumericVector computeAllLogLkhd(NumericVector observedCases,
 //' Compute log likelihood ratio test statistic for all 
 //' 
 //' @inheritParams computeAllLogLkhd
-//' @param permutedCaseMatrix blah
+//' @param permutedCaseMatrix matrix of null simulated cases
 //' @return Max
 //' @export
 // [[Rcpp::export]]

### 2015-10-23 testing.R

@@ -10,30 +10,75 @@ NYleukemia <- SpatialPolygonsDataFrame(sp, data)
 devtools::use_data(NYleukemia, NYleukemia, overwrite=TRUE)
 
 
-
-# Testing define_single_zones
-data("NYleukemia")
-centroids <- coordinates(NYleukemia)
-population <- NYleukemia$population
-pop_upper_bound <- 0.15
-testing <- define_single_zones(centroids, population, pop_upper_bound)
+# Testing plotting function
+options(stringsAsFactors=FALSE)
 
 
+library(maptools)
+shapefile_name <- system.file("extdata/sids/sids.shp", package="SpatialEpi")
+nc_sids <- maptools::readShapePoly(
+  fn = shapefile_name,
+  ID = "FIPSNO",
+  proj4string = CRS("+proj=longlat +ellps=clrk66")
+)
 
-# Testing create_geo_objects
-testing <- create_geo_objects(centroids, population, pop_upper_bound)
+
+library(rgdal)
+shapefile_dir <- "/Library/Frameworks/R.framework/Versions/3.2/Resources/library/spdep/etc/shapes/"
+nc_sids <- readOGR(
+  dsn=shapefile_dir, 
+  layer="sids", 
+  p4s = "+proj=longlat +ellps=clrk66",
+  verbose = FALSE
+)
 
 
 
-# Testing kulldorff
-library(SpatialEpi)
-data("NYleukemia")
-centroids <- coordinates(NYleukemia)
-cases <- NYleukemia$cases
-population <- NYleukemia$population
-expected_cases <- NULL
-pop_upper_bound <- 0.15
-alpha_level <- 0.05
-n_sim <- 9999
-testing <- kulldorff(centroids, cases, population, expected_cases, pop_upper_bound, 
-                     alpha_level=0.05)
+sp_obj <- nc_sids
+variable <- "SID74"
+x_label <- "longitude"
+y_label <- "latitude"
+title <- "North Carolina SIDS Deaths"
+legend_label <- "SIDS Deaths"
+
+plot <- map_variable(sp_obj, "SID74", x_label, y_label)
+
+map_variable <- function(sp_obj, variable=NULL, x_label = NULL, y_label = NULL, 
+                        title = NULL, legend_label = NULL){
+  if(is.null(x_label)){
+    x_label <- "eastings"
+  }
+  if(is.null(y_label)){
+    y_label <- "northings"
+  }
+  if(is.null(legend_label)){
[…]

### 2015-10-23 vignettes/SpatialEpi.R

@@ -2,40 +2,27 @@
 #  library(rgdal)
 #  library(rgeos)
 
-## ---- message=FALSE, collapse=TRUE, tidy=TRUE,tidy.opts=list(width.cutoff=40)----
-library(maptools)
-library(dplyr)
+## ---- message=FALSE, collapse=TRUE, tidy=TRUE, tidy.opts=list(width.cutoff=40), cache=TRUE----
 library(spdep)
-# Awkward
 data(nc.sids)
-sids <- data.frame(
-  Observed=nc.sids$SID74,
-  Expected=nc.sids$BIR74*sum(nc.sids$SID74)/sum(nc.sids$BIR74),
-  x=nc.sids$x, 
-  y=nc.sids$y)
 
-nc.sids <- readShapePoly(
-  fn = system.file("etc/shapes/sids.shp", package="spdep")[1],
-  ID="FIPSNO",
-  proj4string=CRS("+proj=longlat +ellps=clrk66"))
-
-# Create a copy to modify
-nc.sids@data <- nc.sids@data %>% 
-  mutate(
-    Y = SID74,
-    E = sum(Y) * BIR74/sum(BIR74),
-    SMR74 = Y/E,
-    EXP74 = E
+library(rgdal)
+shapefile_dir <- "/Library/Frameworks/R.framework/Versions/3.2/Resources/library/spdep/etc/shapes/"
+nc_sids <- readOGR(
+  dsn=shapefile_dir, 
+  layer="sids", 
+  p4s = "+proj=longlat +ellps=clrk66",
+  verbose = FALSE
   )
-brks <- seq(0, 5)
 
-## ---- echo=TRUE, collapse=TRUE,fig.height=4.5,fig.width=4, fig.cap="Map of SMRs for SIDS in 1974 in North Carolina", tidy.opts=list(width.cutoff=35)----
-spplot(nc.sids, "SMR74", at=brks, col.regions=grey.colors(5,start=.9,end=.1))
-
-## ---- fig.show='hold'----------------------------------------------------
-plot(1:10)
-plot(10:1)
-
-## ---- echo=FALSE, results='asis'-----------------------------------------
-knitr::kable(head(mtcars, 10))
+# Read in shapefile
+options(stringsAsFactors=FALSE)
+library(maptools)
+shapefile_name <- system.file("etc/shapes/sids.shp", package="spdep")
[…]

### 2015-10-23 testing.R
syncing.

@@ -10,75 +10,16 @@ NYleukemia <- SpatialPolygonsDataFrame(sp, data)
 devtools::use_data(NYleukemia, NYleukemia, overwrite=TRUE)
 
 
-# Testing plotting function
+# Two different ways of loading in shapefiles
 options(stringsAsFactors=FALSE)
-
-
 library(maptools)
 shapefile_name <- system.file("extdata/sids/sids.shp", package="SpatialEpi")
-nc_sids <- maptools::readShapePoly(
-  fn = shapefile_name,
-  ID = "FIPSNO",
+nc_sids <- maptools::readShapePoly(fn = shapefile_name, ID = "FIPSNO",
   proj4string = CRS("+proj=longlat +ellps=clrk66")
 )
 
-
 library(rgdal)
 shapefile_dir <- "/Library/Frameworks/R.framework/Versions/3.2/Resources/library/spdep/etc/shapes/"
-nc_sids <- readOGR(
-  dsn=shapefile_dir, 
-  layer="sids", 
-  p4s = "+proj=longlat +ellps=clrk66",
+nc_sids <- rgdal::readOGR(dsn=shapefile_dir, layer="sids", p4s = "+proj=longlat +ellps=clrk66",
   verbose = FALSE
 )
-
-
-
-sp_obj <- nc_sids
-variable <- "SID74"
-x_label <- "longitude"
-y_label <- "latitude"
-title <- "North Carolina SIDS Deaths"
-legend_label <- "SIDS Deaths"
-
-plot <- map_variable(sp_obj, "SID74", x_label, y_label)
-
-map_variable <- function(sp_obj, variable=NULL, x_label = NULL, y_label = NULL, 
-                        title = NULL, legend_label = NULL){
-  if(is.null(x_label)){
-    x_label <- "eastings"
-  }
-  if(is.null(y_label)){
-    y_label <- "northings"
-  }
-  if(is.null(legend_label)){
-    legend_label <- variable
-  }
-  
-  # Create data frame from SpatialPolygonsDataFrame object for ggplot'ing
-  sp_obj@data$id <- rownames(sp_obj@data)
-  map_points <- fortify(sp_obj, region="id")
-  map_df <- inner_join(map_points, sp_obj@data, by="id")
-  
-  # Set up base plot
-  map_ggplot <- 
-    ggplot(map_df, aes(x=long, y=lat, group=group)) +
-    coord_map() +
-    theme_bw() +
-    xlab(x_label) + 
-    ylab(y_label) + 
[…]

### 2015-11-03 vignettes/SpatialEpi.R
Working on Vignettes.

@@ -1,13 +1,0 @@
-## ---- eval=FALSE---------------------------------------------------------
-#  library(rgdal)
-#  library(rgeos)
-
-## ---- message=FALSE, collapse=TRUE, tidy=TRUE, tidy.opts=list(width.cutoff=40), cache=TRUE----
-shapefile_name <- system.file("extdata/sids/sids.shp", package="SpatialEpi")
-proj4string <- sp::CRS("+proj=longlat +ellps=clrk66")
-nc_sids <- maptools::readShapePoly(fn = shapefile_name, ID = "FIPSNO",
-    proj4string = proj4string)
-map_variable(nc_sids)
-names(nc_sids)
-nc_sids$RATE74 <- nc_sids$SID74/nc_sids$BIR74
-

### 2015-11-03 R/mapping.R
Updating gpclib.

@@ -22,6 +22,7 @@ utils::globalVariables(c("lat", "long", "group", ".", "variable"))
 #' @param legend_title label for legend. Defaults to \code{variable_name}.
 #'
 #' @importFrom RColorBrewer brewer.pal
+#' @importFrom gpclib gpclibPermit
 #'
 #' @return a ggplot object
 #' @export
@@ -44,6 +45,8 @@ utils::globalVariables(c("lat", "long", "group", ".", "variable"))
 map_variable <- function(sp_obj, variable_name = NULL, type = "continuous", n_int=5,
                          x_label = "eastings", y_label = "northings", lwd=0.5,
                          title = NULL, legend_title = NULL){
+  gpclibPermit()
+  
   if(is.null(legend_title)){
     legend_title <- variable_name
   }

### 2016-07-08 README.Rmd
Started devtools revamp of SpatialEpi.

--- 2014-06-20 README.Rmd
@@
 ---
-title: "README"
-output: html_document
+title: "SpatialEpi"
+output: github_document
 ---
 
-SpatialEpi: Methods and Data for Spatial Epidemiology
-===========================================
-We load the data and convert the coordinate system from latitude/longitude to a 
-grid-based system.  
-```{r}
-require(SpatialEpi)
-data(NYleukemia)
-map <- NYleukemia$spatial.polygon
-centroids <- latlong2grid(NYleukemia$geo[, 2:3])
-population <- NYleukemia$data$population
-cases <- NYleukemia$data$cases
+[![Travis-CI Build Status](https://travis-ci.org/rudeboybert/SpatialEpi.svg?branch=master)](https://travis-ci.org/rudeboybert/SpatialEpi)
+[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/SpatialEpi)](http://cran.r-project.org/package=SpatialEpi)
+
+Package of data and methods for spatial epidemiology.
+
+## Installation
+
+Get the released version from CRAN:
+
+```R
+install.packages("SpatialEpi")
 ```
 
-There were 4 census tracts that were completely surrounded by another census tract (an enclave like Lesotho).
-We merge them into their surrounding census tracts and modify the population/case counts and SpatialPolygons object accordingly.
-```{r}
-remove <- NYleukemia$surrounded
-add <- NYleukemia$surrounding
-population[add] <- population[add] + population[remove]
-population <- population[-remove]
-cases[add] <- cases[add] + cases[remove]
-cases <- cases[-remove]
-centroids <- centroids[-remove, ]
-map <- SpatialPolygons(map@polygons[-remove], proj4string=CRS("+proj=longlat"))
+Or the development version from GitHub:
+
+```R
+# If you haven't installed devtools yet, do so:
+# install.packages("devtools")
+devtools::install_github("rudeboybert/SpatialEpi")
 ```
-
-We plot the incidence of leukemia for each census tract.  
-```{r, fig.height=8, fig.width=8}
-plotmap(cases/population, map, log=TRUE, nclr=5)
-points(grid2latlong(centroids), pch=4)
-```

### 2016-07-08 R/EBpostdens.R

@@ -1,12 +1,0 @@
-EBpostdens <-
-function(Y, E, alpha, beta, Xrow=NULL, lower=NULL, upper=NULL, main=""){
-  if (is.null(Xrow)) Xrow <- matrix(c(1),nrow=1,ncol=1)
-  xvals <- seq(lower,upper,.01)
-  mu <- as.numeric(exp(Xrow %*% beta))
-  yvals <- dgamma(xvals,alpha+Y,(alpha+E*mu)/mu)
-  plot(xvals,yvals,type="n",
-       xlab=expression(theta),ylab="EB density",cex.lab=1.2)
-  title(paste(main))
-  lines(xvals,yvals)
-  lines(c(Y/E,Y/E),c(0,max(yvals)),lty=2)
-}

### 2016-07-08 R/EBpostthresh.R

@@ -1,7 +1,0 @@
-EBpostthresh <-
-function(Y, E, alpha, beta, Xrow=NULL, rrthresh){
-  if (is.null(Xrow)) Xrow <- matrix(rep(1,length(Y)),nrow=length(Y),ncol=1)
-  mu <- as.numeric(exp(Xrow %*% beta))
-  thresh <- 1-pgamma(rrthresh,alpha+Y,(alpha+E*mu)/mu)
-  return(thresh=thresh)
-}

### 2016-07-08 R/GammaPriorCh.R

@@ -1,7 +1,0 @@
-GammaPriorCh <-
-function(theta, prob, d){
-	a <- d/2
-	b <- 0.5*2*a*theta^2/qt(p=prob,df=2*a)^2
-	cat("Gamma Parameters: ",a,b,"\n")
-	list(a=a,b=b)
-}

### 2016-07-08 R/LogNormalPriorCh.R

@@ -1,8 +1,0 @@
-LogNormalPriorCh <-
-function(theta1, theta2, prob1, prob2){	
-zq1 <- qnorm(prob1)
-zq2 <- qnorm(prob2)
-mu <- log(theta1)*zq2/(zq2-zq1) - log(theta2)*zq1/(zq2-zq1)
-sigma <- (log(theta1)-log(theta2))/(zq1-zq2)
-list(mu=mu,sigma=sigma)
-}

### 2016-07-08 R/SpatialEpi.R

@@ -1,0 +1,13 @@
+#' SpatialEpi: Data and methods for spatial epidemiology.
+#'
+#' This package provides...
+#' 
+#' @section Foo functions:
+#' The foo functions ...
+#'
+#' @docType package
+#' @name SpatialEpi
+#' @useDynLib SpatialEpi
+#' @importFrom Rcpp sourceCpp
+NULL
+#> NULL

### 2016-07-08 R/bayes_cluster.R

@@ -1,105 +1,0 @@
-bayes_cluster <-
-function(y, E, population, sp.obj, centroids, max.prop, shape, rate, J, pi0,
-         n.sim.lambda, n.sim.prior, n.sim.post, burnin.prop = 0.1,
-         theta.init = vector(mode="numeric", length=0)
-){
-print(paste("Algorithm started on:", date()))
-  
-# MCMC parameters
-pattern <- c(0, 1)
-p.moves <- normalize(c(1, 1, 1, 1, 1))
-  
-# Geographic info and create geographical objects to use
-geo.objects <- create_geo_objects(max.prop, population, centroids, sp.obj)
-overlap <- geo.objects$overlap
-cluster.coords <- geo.objects$cluster.coords    
-n <- length(E)
-n.zones <- nrow(cluster.coords)
-
-# Set cutoffs of the relative risk to determine low and high risk
-RR.range <- seq(0.7, 1/0.7, length=100000)
-pN <- dgamma(RR.range, shape[1], rate[1])
-pW <- dgamma(RR.range, shape[2], rate[2])
-cutoffs <- list(
-  low=RR.range[min(which(pN/pW>1))],
-  high=RR.range[max(which(pN/pW>1))]
-)
-
-
-#-------------------------------------------------------------------------------
-# Obtain values of lambda using importance sampling
-#-------------------------------------------------------------------------------
-# Set uniform prior on single zones.
-prior.z <- rep(1/n.zones, n.zones)
-log_prior.z <- log(prior.z) - log(sum(prior.z))
-
-# Estimate q and lambda
-results <- estimate_lambda(n.sim.lambda, J, prior.z, overlap, pi0)
-lambda <- results$lambda
-prior.j <- results$prior.j
-print(paste("Importance sampling of lambda complete on:", date()))
-
-
-#-------------------------------------------------------------------------------
-# Obtain prior map
-#-------------------------------------------------------------------------------
-# Generate MCMC samples from prior
-prior.chain <- MCMC_simulation(n.sim.prior, pattern, theta.init, overlap, 
-                               cluster.coords, p.moves, J, prior.z, lambda)
-
-# Trim burn-in
-burnin <- n.sim.prior * burnin.prop
-prior.sample <- prior.chain$sample[-c(1:burnin)]
-
-# Prior Probs of Cluster Membership for each Area
-param.prior.zone <- list(shape=rep(shape[2],n.zones), rate=rep(rate[2],n.zones))
[…]

### 2016-07-08 R/circle.R

@@ -1,17 +1,0 @@
-circle <-
-function(geo, cluster.center, cluster.end){
-
-# Compute interpoint distance
-distance <- dist(as.matrix(geo), upper=TRUE, diag=TRUE)
-distance <- as.matrix(distance)[cluster.center, cluster.end]
-
-# For drawing radius of cluster on map
-polar <- seq(0, 2*pi, length=1000)
-        
-cluster.radius <- data.frame(cbind(
-  x=distance*cos(polar)+geo$x[cluster.center],
-  y=distance*sin(polar)+geo$y[cluster.center]
-  ))
-
-return(cluster.radius)
-}

### 2016-07-08 R/create_geo_objects.R

@@ -1,81 +1,0 @@
-create_geo_objects <-
-function(max.prop, population, centroids, sp.obj){
-
-# Number of areas
-n <- nrow(centroids)
-
-
-#-----------------------------------------------------------------------------
-# Set up single zones
-#-----------------------------------------------------------------------------
-if(max.prop < max(normalize(population))){
-  print(paste("max.prop needs to be at least", max(normalize(population))))
-}
-
-# Get basic single zone info
-geoInfo <- zones(centroids, population, max.prop)
-nearest.neighbors <- geoInfo$nearest.neighbors
-cluster.coords <- geoInfo$cluster.coords
-n.zones <- nrow(cluster.coords)
-
-# Create list of length n.zones indicating the component areas for each zone
-cluster.list <- vector(mode="list", length=n.zones)
-counter <- 1
-for(i in 1:n) {
-  nn <- nearest.neighbors[[i]]
-  for(j in 1:length(nn)) {
-    cluster.list[[counter]] <- nn[1:j] 
-    counter <- counter + 1  
-  } 
-}
-
-#-----------------------------------------------------------------------------
-# Generate overlap object which tracks the overlap between single zones
-#-----------------------------------------------------------------------------
-# For each area, get "presense":  list all unit zones that it is included in
-presence <- vector(mode="list",length=n)
-for(i in 1:n){
-  presence[[i]] <- which(sapply(cluster.list, function(x){is.element(i, x)}))
-}
-
-
-#-------------------------------------------------------------------------------
-# Compute "buffer" of areas between single zones.
-#-------------------------------------------------------------------------------
-# Compute adjacency matrix
-adj <- poly2nb(sp.obj, queen=TRUE)
-adj <- nb2mat(adj, zero.policy=TRUE)[1:n, 1:n]
-
-# Convert Adjacency Matrix To A List
-adj.new <- vector("list",length=n)
-for(i in 1:n){
-  adj.new[[i]] <- which(adj[i,]!=0)
-}
-adj <- adj.new
-rm(adj.new)
-
-# Update presence list to incorporate buffer.  We need to preserve original 
-# presence list all the way thru, so temporarily store results here
-presence_temp <- presence
-
-# Loop thru all areas
-for(i in 1:n){
-  current.zones <- presence[[i]]
-  adjacent.areas <- adj[[i]]
-  adjacent.zones <- unique(unlist(presence[adjacent.areas]))
[…]

### 2016-07-08 R/data.R

@@ -1,0 +1,60 @@
+#' Leukemia in Upstate New York.
+#'
+#' Leukemia data for 281 census tracts in 7 counties in Upstate New York.
+#'
+#' @format object of class \code{\link[sp]{SpatialPolygonsDataFrame-class}} with a data frame of
+#' 281 observations and 4 variables:
+#' \itemize{
+#'   \item ID: id variable
+#'   \item FIPS: Federal Information Processing Standards code of census tract
+#'   \item population: population size (1980 U.S. Census)
+#'   \item cases: number of cases 1978-1982
+#' }
+#' @references Waller, L. and C. Gotway (2004) Applied Spatial Statistics for
+#' Public Health Data. New York: John Wiley and Sons.
+#' @source \url{http://web1.sph.emory.edu/users/lwaller/ch9index.htm}
+"NYleukemia"
+
+
+#' Pennsylvania Lung Cancer.
+#'
+#' County-level (n=67) population/case data for lung cancer in Pennsylvania in 
+#' 2002, stratified on race (white vs non-white), gender and age (Under 40, 
+#' 40-59, 60-69 and 70+).  Additionally, county-specific smoking rates. 
+#'
+#' @format object of class \code{\link[sp]{SpatialPolygonsDataFrame-class}} with a data frame of
+#' 281 observations and 4 variables:
+#' \itemize{
+#'   \item ID: id variable
+#'   \item FIPS: Federal Information Processing Standards code of census tract
+#'   \item population: population size (1980 U.S. Census)
+#'   \item cases: number of cases 1978-1982
+#' }
+#' @source Population data was obtained from the 2000 decennial census, lung 
+#' cancer and smoking data were obtained from the Pennsylvania Department of 
+#' Health website: \url{http://www.dsf.health.state.pa.us/}
+"pennLC"
+
+
+#' Lip Cancer in Scotland.
+#'
+#' County-level (n=56) data for lip cancer among males in Scotland between 
+#' 1975-1980.
+#'
+#' @format object of class \code{\link[sp]{SpatialPolygonsDataFrame-class}} with a data frame of
+#' 281 observations and 4 variables:
+#' \itemize{
+#'   \item ID: id variable
+#'   \item FIPS: Federal Information Processing Standards code of census tract
+#'   \item population: population size (1980 U.S. Census)
+#'   \item cases: number of cases 1978-1982
+#' }
[…]

### 2016-07-08 R/eBayes.R

@@ -1,21 +1,0 @@
-eBayes <-
-function(Y, E, Xmat=NULL){
-	# Check for covariates
-	if (is.null(Xmat)) {
-		mod <- glm.nb(Y ~ 1+offset(log(E)), link=log)
-	}else{
-		mod <- glm.nb(Y ~ Xmat+offset(log(E)), link=log)
-	}
-
-	# Get MLE's of parameters
-	alpha <- mod$theta
-	muhat <- mod$fitted/E
-	
-	# Generate weights between global and local SMR
-	wgt <- E*muhat/(alpha+E*muhat); SMR <- Y/E
-	RR <- as.numeric(wgt*SMR + (1-wgt)*muhat)
-	RRmed <- qgamma(0.5,alpha+Y,(alpha+E*muhat)/muhat)
-	
-	# Output results
-	list(RR=RR, RRmed=RRmed, beta=mod$coeff, alpha=alpha, SMR=SMR)
-}

### 2016-07-08 R/estimate_lambda.R

@@ -1,44 +1,0 @@
-estimate_lambda <-
-function(n.sim, J, prior.z, overlap, pi0){
-
-n.zones <- length(prior.z)
-
-# q_0 = q_1 = 1, since there is no concept of overlap for j=0, 1
-q <- c(1, 1, rep(0, J-1))
-var.q <- c(NA, NA, rep(0, J-1))
-
-
-#-------------------------------------------------------------------------------
-# All k's, sample configurations of k single zones, and see if the k single
-# zones overlap or not
-#-------------------------------------------------------------------------------
-for(k in 2:J){
-  unit.zones <- sample(1:n.zones, k*n.sim, replace=TRUE, prob=prior.z)
-  unit.zones <- matrix(unit.zones, ncol=k)
-  no.overlap <- check_overlap(unit.zones, overlap)
-  
-  denom <- rep(factorial(k), n.sim)
-  
-  # Compute multinomial coefficient watching out for when certain single zones
-  # are sampled more than once
-  duplicates <- apply(unit.zones, 1, function(x){length(unique(x))})
-  duplicates <- which(duplicates != k)
-  for(i in duplicates) {
-    denom[i] <-  factorial(k)/prod(factorial(table(unit.zones[i,])))
-  }
-  
-  q[k+1] <- mean(no.overlap/denom)
-  var.q[k+1] <- var(no.overlap/denom)/n.sim
-}
-
-
-#-------------------------------------------------------------------------------
-# Using q compute estimate of lambda and prior probability of j clusters
-#-------------------------------------------------------------------------------
-lambda <- (1-pi0)/( (1-pi0)*J + pi0*sum(q[-1]) )
-lambda <- rep(lambda, J)
-lambda <- c(1-sum(lambda), lambda)
-prior.j <- normalize(lambda*q)
-
-return(list(lambda=lambda, prior.j=prior.j))
-}

### 2016-07-08 R/kulldorff.R

@@ -1,143 +1,0 @@
-kulldorff <-
-function(geo, cases, population, expected.cases=NULL, pop.upper.bound, 
-         n.simulations, alpha.level, plot=TRUE){
-
-#-------------------------------------------------------------------------------
-# Initialization 
-#-------------------------------------------------------------------------------
-# Determine likelihood type: binomial or poisson
-if(is.null(expected.cases)){
-	type <- "binomial"
-	denominator <- population
-	expected.cases <- sum(cases) * (denominator/sum(denominator))
-# poisson case	
-}else{
-	type <- "poisson"
-	denominator <- expected.cases
-}
-
-# Get geographic information
-geo.results <- zones(geo, population, pop.upper.bound)
-nearest.neighbors <- geo.results$nearest.neighbors
-cluster.coords <- geo.results$cluster.coords
-n.zones <- nrow(cluster.coords)
-
-
-#-------------------------------------------------------------------------------
-# Observed statistic computation
-#-------------------------------------------------------------------------------
-lkhd <- computeAllLogLkhd(cases, denominator, nearest.neighbors, n.zones, type)
-
-# Get areas included in most likely cluster
-cluster.index <- which.max(lkhd)
-
-# cluster center and radial area
-center <- cluster.coords[cluster.index,1]
-end <- cluster.coords[cluster.index,2]
-
-# list of all areas included in cluster	
-cluster <- nearest.neighbors[[center]]
-cluster <- cluster[1:which(cluster == end)]
-
-
-#-------------------------------------------------------------------------------
-# Compute Monte Carlo randomized p-value
-#-------------------------------------------------------------------------------
-# Simulate cases under null hypothesis of no area effects i.e. conditioned on E
-perm <- rmultinom(n.simulations, round(sum(cases)), prob=denominator)
-
-# Compute simulated lambda's:  max log-lkhd in region
-sim.lambda <- kulldorffMC(perm, denominator, nearest.neighbors, n.zones, type)
-
-# Compute Monte Carlo p-value
-combined.lambda <- c(sim.lambda, max(lkhd))
-p.value <- 1-mean(combined.lambda < max(lkhd))
-
-# Plot histogram
-if(plot){
-	hist(combined.lambda, 
-       main="Monte Carlo Distribution of Lambda",
-       xlab=expression(log(lambda)))
-	abline(v=max(lkhd), col="red")
-	legend("top",
[…]

### 2016-07-08 R/latlong2grid.R

@@ -1,50 +1,0 @@
-latlong2grid <-
-function(input){
-  
-toradians <- atan(1)/45
-radiusearth <- 0.5*(6378.2+6356.7)
-sine51 <- sin( 51.5*toradians )
-
-
-#-------------------------------------------------------------------------------
-# If a Spatial Polygon
-#-------------------------------------------------------------------------------
-if(is(input)[1] == "SpatialPolygons"){  
-  for( i in 1:length(input@polygons) ){
-    # for all Polygons's in polygon
-    for( j in 1:length(input@polygons[[i]]@Polygons) ){
-      # Convert coordinates
-      new.coords <- cbind(
-        (input@polygons[[i]]@Polygons[[j]]@coords[,1]*toradians)*radiusearth*sine51,
-        (input@polygons[[i]]@Polygons[[j]]@coords[,2]*toradians)*radiusearth
-      )
-      new.coords <- as.matrix(new.coords)
-      colnames(new.coords) <- NULL
-      rownames(new.coords) <- NULL
-      
-      # Update Polygons
-      input@polygons[[i]]@Polygons[[j]]@coords <- new.coords
-      input@polygons[[i]]@Polygons[[j]] <- Polygon(input@polygons[[i]]@Polygons[[j]])
-    }
-    # Update polygons
-    input@polygons[[i]] <- Polygons(
-      input@polygons[[i]]@Polygons,
-      ID=input@polygons[[i]]@ID
-    )	
-  }
-  
-  output <- SpatialPolygons(input@polygons,proj4string=CRS("+proj=utm"))
-
-  
-#-------------------------------------------------------------------------------  
-# else return numeric
-#-------------------------------------------------------------------------------
-}else{
-  output <- data.frame(cbind(
-    x=(input[,1]*toradians)*radiusearth*sine51,
-    y=(input[,2]*toradians)*radiusearth
-  ))		
-}
-
-return(output)	
-}

### 2016-07-08 R/mapvariable.R

@@ -1,88 +1,0 @@
-mapvariable <-
-function(y, spatial.polygon, ncut=1000, nlevels=10, lower=NULL, upper=NULL, 
-         main=NULL, xlab=NULL, ylab=NULL){
-
-#-------------------------------------------------------------------------------
-# Create id indicators for coloring scheme
-#-------------------------------------------------------------------------------
-if (is.null(lower)) 
-	lower <- min(y) #- 0.0001
-if (is.null(upper)) 
-	upper <- max(y) #+ 0.0001
-
-id <- cut(y, breaks=seq(from=lower, to=upper, length=(ncut+1)))
-id <- as.numeric(id)
-id[is.na(id)] <- 0
-id <- id + 1
-
-# Set colours to grey scale
-palette(gray(seq(1,0, len=(ncut+1))))
-
-
-#-------------------------------------------------------------------------------
-# Make the scale of the two axes the same
-#-------------------------------------------------------------------------------
-xrnge <- spatial.polygon@bbox[1,]
-yrnge <- spatial.polygon@bbox[2,]	
-
-xd <- xrnge[2] - xrnge[1]
-yd <- yrnge[2] - yrnge[1]
-
-if(xd > yd){		
-	xplot <- xrnge
-	yplot <- NULL
-	yplot[1] <- ((yrnge[2] + yrnge[1])/2) - xd/2
-	yplot[2] <- ((yrnge[2] + yrnge[1])/2) + xd/2
-}
-
-if(xd <= yd){
-	yplot <- yrnge
-	xplot <- NULL
-	xplot[1] <- ((xrnge[2] + xrnge[1])/2) - yd/2
-	xplot[2] <- ((xrnge[2] + xrnge[1])/2) + yd/2
-}
-
-
-#-------------------------------------------------------------------------------
-# Plots
-#-------------------------------------------------------------------------------
-def.par <- par(no.readonly = TRUE)
-layout(matrix(c(1,2),ncol=2,nrow=1), heights=c(.3,.3), widths=c(.4,.1))
-
-# plot variable
-plot(spatial.polygon, axes=TRUE, col=id)
-
-# plot title and axis labels
-if(!is.null(main)){
-	title(main=main)
-}
-if(!is.null(xlab)){
[…]

### 2016-07-08 R/plotmap.R

@@ -1,50 +1,0 @@
-plotmap <-
-function(values, map, log=FALSE, nclr=7, include.legend=TRUE, lwd=0.5, round=3,
-         brks=NULL, legend=NULL, location='topright', rev=FALSE){
-
-# create colors, each based on quantiles of data
-plotclr <- grey(1-seq(0, 1, by=1/(nclr-1)))
-
-nclr <- nclr + 1
-if(is.null(brks)){
-  if(log){
-    brks <- exp(
-      seq(from=min(log(values)), to=max(log(values)), length.out=nclr)
-      )
-  }else{
-    brks <- seq(from=min(values), to=max(values), length.out=nclr)
-  }
-}
-nclr <- nclr - 1
-
-
-# Assign colors based on intervals
-if(rev){
-  plotclr <- rev(plotclr)
-}
-colornum <- findInterval(values, brks, all.inside=T);
-colcode <- plotclr[colornum]
-
-
-# Modify legend
-if(is.null(legend)){
-  legend <- leglabs(signif(brks,digits=round))
-  # First element of legend
-  legend[1] <- paste(
-    signif(min(values), digits=round), "-", 
-    substr(legend[1], 7, nchar(legend[1]))
-  )
-  # Last element of legend
-  legend[nclr]<- paste(
-    substr(legend[nclr], 6, nchar(legend[nclr])), "-",
-    signif(max(values), digits=round)
-  )
-}
-
-
-plot(map, axes=TRUE, col=colcode, lwd=lwd)
-if(include.legend) {
-  legend(location, legend=legend, fill=plotclr, bty="n")
-}
-
-}

### 2016-07-08 R/zones.R

@@ -1,55 +1,0 @@
-zones <-
-function(geo, population, pop.upper.bound){
-
-# number of areas
-n <- nrow(geo)
-# total population
-total.pop <- sum(population)
-# Interpoint distance matrix
-dist <- as.matrix(dist(as.matrix(geo), upper=TRUE, diag=TRUE))
-
-
-#-------------------------------------------------------------------------------
-# For each area, list of closest neighbors up until pop.upper.bound of
-# population is met.  We count the number of candidate zones in n.zones 
-# Note:  for each county, the list of nearest neighbors form the total number of
-# candidate zones
-#-------------------------------------------------------------------------------
-nearest.neighbors <- vector(mode="list", length=n)
-
-n.zones <- 0
-vector.cutoffs <- rep(0, n+1)
-
-for(i in 1:n) {	
-  # Sort the areas by distance, then include them one-by-one until cluster bound
-  # is met
-  neighbors <- order(dist[,i])
-  
-  # include only up until pop.upper.bound is hit
-  neighbors <- neighbors[ which( 
-      cumsum(population[neighbors])/total.pop <= pop.upper.bound
-    )] 
-  
-  nearest.neighbors[[i]] <- neighbors
-  
-  # Update total number of zones
-  n.zones <- n.zones + length(neighbors)
-  
-  # internal
-  vector.cutoffs[i+1] <- n.zones	
-}
-
-cluster.coords <- cbind(rep(1:n, times=diff(vector.cutoffs)), 
-                        unlist(nearest.neighbors))
-
-
-#-------------------------------------------------------------------------------
-# Output results
-#-------------------------------------------------------------------------------
-results <- list(
-  nearest.neighbors = nearest.neighbors, 
-  cluster.coords = cluster.coords, 
-  dist=dist)
-
-return(results)
-}

### 2016-07-14 README.Rmd
About to start adding functions.

@@
 
-```R
+```{r, eval=FALSE}
 install.packages("SpatialEpi")
@@
 
-```R
+```{r, eval=FALSE}
 # If you haven't installed devtools yet, do so:
@@
 ```
+
+## Example
+
+We load the data and convert the coordinate system from latitude/longitude to a 
+grid-based system.
+
+```{r, message=FALSE, warning=FALSE}
+library(SpatialEpi)
+```
+
+```{r, message=FALSE, warning=FALSE, eval=FALSE}
+data(NYleukemia)
+sp.obj <- NYleukemia$spatial.polygon
+centroids <- latlong2grid(NYleukemia$geo[, 2:3])
+population <- NYleukemia$data$population
+cases <- NYleukemia$data$cases
+```
+
+We plot the incidence of leukemia for each census tract.
+
+```{r, message=FALSE, warning=FALSE, eval=FALSE}
+plotmap(cases/population, sp.obj, log=TRUE, nclr=5)
+points(grid2latlong(centroids), pch=4)
+```
+
+We run the Bayesian Cluster Detection method from Wakefield and Kim (2013)
+
+```{r, message=FALSE, warning=FALSE, eval=FALSE}
+y <- cases
+E <- expected(population, cases, 1)
+max.prop <- 0.15
+shape <- c(2976.3, 2.31)
+rate <- c(2977.3, 1.31)
+J <- 7
+pi0 <- 0.95
+n.sim.lambda <- 10^4
+n.sim.prior <- 10^5
+n.sim.post <- 10^5
+
+# Compute output
+output <- bayes_cluster(y, E, population, sp.obj, centroids, max.prop, 
+ shape, rate, J, pi0, n.sim.lambda, n.sim.prior, n.sim.post)
+```
+
+```{r, eval=FALSE}
+plotmap(output$post.map$high.area, sp.obj)
+```

### 2016-08-19 README.Rmd
Starting clean-up.

@@
 ---
-title: "SpatialEpi"
 output: github_document
@@
 
+<!-- README.md is generated from README.Rmd. Please edit that file -->
+
+```{r, echo = FALSE}
+knitr::opts_chunk$set(
+  collapse = TRUE,
+  comment = "#>",
+  fig.path = "README-"
+)
+```
+
+# SpatialEpi
+
 [![Travis-CI Build Status](https://travis-ci.org/rudeboybert/SpatialEpi.svg?branch=master)](https://travis-ci.org/rudeboybert/SpatialEpi)
 [![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/SpatialEpi)](http://cran.r-project.org/package=SpatialEpi)
+[![CRAN RStudio mirror downloads](http://cranlogs.r-pkg.org/badges/SpatialEpi)](http://www.r-pkg.org/pkg/SpatialEpi)
 

### 2016-08-19 README.Rmd
Done first pass of refactoring bayes_cluster code.

@@
   comment = "#>",
-  fig.path = "README-"
+  fig.path = "README_figure/README-"
 )
@@
 
-```{r, message=FALSE, warning=FALSE, eval=FALSE}
+```{r, message=FALSE, warning=FALSE}
 data(NYleukemia)
@@
 
-```{r, message=FALSE, warning=FALSE, eval=FALSE}
+```{r, message=FALSE, warning=FALSE}
 plotmap(cases/population, sp.obj, log=TRUE, nclr=5)
@@
 
-```{r, message=FALSE, warning=FALSE, eval=FALSE}
+```{r bayes_cluster, message=FALSE, warning=FALSE, cache=TRUE}
 y <- cases
@@
 # Compute output
-output <- bayes_cluster(y, E, population, sp.obj, centroids, max.prop, 
- shape, rate, J, pi0, n.sim.lambda, n.sim.prior, n.sim.post)
+# output <- bayes_cluster(y, E, population, sp.obj, centroids, max.prop,
+#                         shape, rate, J, pi0, n.sim.lambda, n.sim.prior,
+#                         n.sim.post)
 ```
 
-```{r, eval=FALSE}
+```{r, cache=TRUE}
 plotmap(output$post.map$high.area, sp.obj)

### 2016-08-19 R/bayes_cluster.R

@@ -1,259 +1,299 @@
-bayes_cluster <-
-  function(y, E, population, sp.obj, centroids, max.prop, shape, rate, J, pi0,
-           n.sim.lambda, n.sim.prior, n.sim.post, burnin.prop = 0.1,
-           theta.init = vector(mode="numeric", length=0)
-  ){
-    print(paste("Algorithm started on:", date()))
-    
-    # MCMC parameters
-    pattern <- c(0, 1)
-    p.moves <- normalize(c(1, 1, 1, 1, 1))
-    
-    # Geographic info and create geographical objects to use
-    geo.objects <- create_geo_objects(max.prop, population, centroids, sp.obj)
-    overlap <- geo.objects$overlap
-    cluster.coords <- geo.objects$cluster.coords    
-    n <- length(E)
-    n.zones <- nrow(cluster.coords)
-    
-    # Set cutoffs of the relative risk to determine low and high risk
-    RR.range <- seq(0.7, 1/0.7, length=100000)
-    pN <- dgamma(RR.range, shape[1], rate[1])
-    pW <- dgamma(RR.range, shape[2], rate[2])
-    cutoffs <- list(
-      low=RR.range[min(which(pN/pW>1))],
-      high=RR.range[max(which(pN/pW>1))]
-    )
-    
-    
-    #-------------------------------------------------------------------------------
-    # Obtain values of lambda using importance sampling
-    #-------------------------------------------------------------------------------
-    # Set uniform prior on single zones.
-    prior.z <- rep(1/n.zones, n.zones)
-    log_prior.z <- log(prior.z) - log(sum(prior.z))
-    
-    # Estimate q and lambda
-    results <- estimate_lambda(n.sim.lambda, J, prior.z, overlap, pi0)
-    lambda <- results$lambda
-    prior.j <- results$prior.j
-    print(paste("Importance sampling of lambda complete on:", date()))
-    
-    
-    #-------------------------------------------------------------------------------
-    # Obtain prior map
-    #-------------------------------------------------------------------------------
-    # Generate MCMC samples from prior
-    prior.chain <- MCMC_simulation(n.sim.prior, pattern, theta.init, overlap, 
-                                   cluster.coords, p.moves, J, prior.z, lambda)
-    
-    # Trim burn-in
-    burnin <- n.sim.prior * burnin.prop
[…]

### 2016-08-19 R/geographic_objects.R

@@ -1,348 +1,468 @@
-circle <-
-  function(geo, cluster.center, cluster.end){
+# R CMD CHECK returns NOTE that there is "no visible binding 
+# for global variable" for variables used by dplyr/ggplot2, hence this hack fix.
+utils::globalVariables(c("distance", "prop_pop", "cum_prop", "."))
+
+
+
+#' Define a study region's single zones.
+#' 
+#' Using the centroids of the n areas in a study region, define the set of 
+#' single zones based on the upper bound \code{pop_upper_bound} of the 
+#' proportion of the study region's population each single zone can contain.
+#' 
+#' @param centroids data frame with 2 columns of the centroids of the n areas 
+#'   in the study region.
+#' @param population vector of length n of the corresponding population 
+#'   counts.
+#' @param pop_upper_bound upper bound of the proportion of the study 
+#'   region's population the single zones can contain.
+#'   
+#' @return A list with two objects \describe{ 
+#'   \item{\code{nearest_neighbors}}{A list of length n of the single zones 
+#'   for each area} 
+#'   \item{\code{cluster_coords}}{A data frame with 2 columns of the centering 
+#'   and radial area for each single zone.} 
+#'   }
+#' @references Kulldorff, M. (1997) A spatial scan statistic. 
+#'   \emph{Communications in Statistics: Theory and Methods}, \bold{26}, 
+#'   1481--1496.
+#' @seealso \code{\link{create_geo_objects}}
+#' @export
+#' @importFrom magrittr %>%
+#'   
+#' @examples 
+#' data(NYleukemia)
+#' centroids <- sp::coordinates(NYleukemia$spatial.polygon)
+#' single_zones <- define_single_zones(centroids, NYleukemia$data$population, 0.15)
+define_single_zones <- function(centroids, population, pop_upper_bound) {
+  # Number of areas
+  n <- nrow(centroids)
+  # Interpoint distance matrix
+  dist_matrix <- 
+    centroids %>% 
+    dist(upper = TRUE, diag = TRUE) %>% 
+    as.matrix()
+  
+  # Area indices and proportion of the population each area contains
+  areas <- data.frame(
+    area = 1:n,
+    prop_pop= population/sum(population)
+  )
+  
+  # For each area, we compute the list of neighbors in order of distance such
+  # that the proportion of the study population that is included is less than or
[…]

### 2016-08-19 R/mapping.R

@@ -1,141 +1,176 @@
-mapvariable <-
-  function(y, spatial.polygon, ncut=1000, nlevels=10, lower=NULL, upper=NULL, 
-           main=NULL, xlab=NULL, ylab=NULL){
-    
-    #-------------------------------------------------------------------------------
-    # Create id indicators for coloring scheme
-    #-------------------------------------------------------------------------------
-    if (is.null(lower)) 
-      lower <- min(y) #- 0.0001
-    if (is.null(upper)) 
-      upper <- max(y) #+ 0.0001
-    
-    id <- cut(y, breaks=seq(from=lower, to=upper, length=(ncut+1)))
-    id <- as.numeric(id)
-    id[is.na(id)] <- 0
-    id <- id + 1
-    
-    # Set colours to grey scale
-    palette(gray(seq(1,0, len=(ncut+1))))
-    
-    
-    #-------------------------------------------------------------------------------
-    # Make the scale of the two axes the same
-    #-------------------------------------------------------------------------------
-    xrnge <- spatial.polygon@bbox[1,]
-    yrnge <- spatial.polygon@bbox[2,]	
-    
-    xd <- xrnge[2] - xrnge[1]
-    yd <- yrnge[2] - yrnge[1]
-    
-    if(xd > yd){		
-      xplot <- xrnge
-      yplot <- NULL
-      yplot[1] <- ((yrnge[2] + yrnge[1])/2) - xd/2
-      yplot[2] <- ((yrnge[2] + yrnge[1])/2) + xd/2
-    }
-    
-    if(xd <= yd){
-      yplot <- yrnge
-      xplot <- NULL
-      xplot[1] <- ((xrnge[2] + xrnge[1])/2) - yd/2
-      xplot[2] <- ((xrnge[2] + xrnge[1])/2) + yd/2
-    }
-    
-    
-    #-------------------------------------------------------------------------------
-    # Plots
-    #-------------------------------------------------------------------------------
-    def.par <- par(no.readonly = TRUE)
-    layout(matrix(c(1,2),ncol=2,nrow=1), heights=c(.3,.3), widths=c(.4,.1))
-    
-    # plot variable
-    plot(spatial.polygon, axes=TRUE, col=id)
-    
[…]

### 2016-08-19 R/other_cluster.R

@@ -1,3 +1,16 @@
+#' Title
+#'
+#' @param geo 
+#' @param population 
+#' @param cases 
+#' @param expected.cases 
+#' @param k 
+#' @param alpha.level 
+#'
+#' @return
+#' @export
+#'
+#' @examples
 besag_newell <-
   function(geo, population, cases, expected.cases=NULL, k, alpha.level){
     
@@ -78,6 +91,16 @@ besag_newell <-
 
 
 
+#' Title
+#'
+#' @param population 
+#' @param cases 
+#' @param n.strata 
+#'
+#' @return
+#' @export
+#'
+#' @examples
 expected <-
   function(population, cases, n.strata){
     
@@ -108,6 +131,21 @@ expected <-
 
 
 
+#' Title
+#'
+#' @param geo 
+#' @param cases 
+#' @param population 
+#' @param expected.cases 
+#' @param pop.upper.bound 
+#' @param n.simulations 
+#' @param alpha.level 
+#' @param plot 
+#'
+#' @return
+#' @export
+#'
+#' @examples
 kulldorff <-
   function(geo, cases, population, expected.cases=NULL, pop.upper.bound, 
            n.simulations, alpha.level, plot=TRUE){

### 2016-08-19 src/bayes_cluster.cpp

@@ -3,7 +3,6 @@
 using namespace Rcpp;
 
 
-
 // [[Rcpp::export]]
 NumericVector normalize(NumericVector x) {
   int n = x.size();
@@ -132,8 +131,8 @@ int ProbSampleReplace(NumericVector prob) {
 	}
 	
 	RNGScope scope;
-	int sample = RcppArmadillo::sample(x, 1, FALSE, prob)[0];  
-	return sample;
+	int sample_vector = RcppArmadillo::sample(x, 1, FALSE, prob)[0];  
+	return sample_vector;
 }
 
 
@@ -141,7 +140,7 @@ int ProbSampleReplace(NumericVector prob) {
 // [[Rcpp::export]]
 NumericVector check_overlap(NumericMatrix config, List overlap) {
   List presence = as<List>(overlap["presence"]);
-  List cluster_list = as<List>(overlap["cluster.list"]);
+  List cluster_list = as<List>(overlap["cluster_list"]);
   int nSim = config.nrow();
   int k = config.ncol(), index, index2;
   int nZones = cluster_list.size();
@@ -257,7 +256,7 @@ NumericMatrix return_death_moves(NumericVector theta) {
 // [[Rcpp::export]]
 NumericMatrix return_birth_moves(NumericVector theta, List overlap) {
 	List presence = as<List>(overlap["presence"]);
-	List cluster_list = as<List>(overlap["cluster.list"]);
+	List cluster_list = as<List>(overlap["cluster_list"]);
 
 	int k = theta.size(), n_zones = cluster_list.size(), counter, zone, zone_area, 
 	n_birth_moves;//, move_possible;
@@ -354,7 +353,7 @@ List return_local_moves(NumericVector theta, List overlap, NumericMatrix
 	cluster_coords) {
 
 	List presence = as<List>(overlap["presence"]);
-	List cluster_list = as<List>(overlap["cluster.list"]);
+	List cluster_list = as<List>(overlap["cluster_list"]);
 
 	int k = theta.size(), n_zones = cluster_list.size(), zone, start, end, zone_sub,
 	zone_sub2, center;
@@ -487,7 +486,7 @@ List MCMC_simulation(int n_sim, NumericVector pattern, NumericVector theta_init,
 
 	// Store Outputs Here
 	NumericVector move_trace(n_sim), accpt_trace(n_sim), ratio(n_sim);
-	List sample(n_sim);
+	List sample_vector(n_sim);
 
[…]

### 2016-08-19 testing/geographic_objects.R

@@ -31,8 +31,8 @@ utils::globalVariables(c("distance", "prop_pop", "cum_prop", "."))
 #'   
 #' @examples 
 #' data(NYleukemia)
-#' centroids <- sp::coordinates(NYleukemia)
-#' single_zones <- define_single_zones(centroids, NYleukemia$population, 0.15)
+#' centroids <- sp::coordinates(NYleukemia$spatial.polygon)
+#' single_zones <- define_single_zones(centroids, NYleukemia$data$population, 0.15)
 define_single_zones <- function(centroids, population, pop_upper_bound) {
   
   # Number of areas

### 2016-08-19 testing/testing.R

@@ -1,19 +1,27 @@
-# TODO
-# * Write better package description
-# * Add URL to Journal of Statistical Software!
-# * Add county names to NYleukemia
-# * Cite SIDS data
-# * Add Scotland data
-# * NYleukemia, scotland source
-# * mapproj, gpclib packages
-# * Fix gpclibPermit()
+# Testing bayes_cluster
+data(NYleukemia)
+sp.obj <- NYleukemia$spatial.polygon
+centroids <- latlong2grid(NYleukemia$geo[, 2:3])
+population <- NYleukemia$data$population
+cases <- NYleukemia$data$cases
+y <- cases
+E <- expected(population, cases, 1)
+max.prop <- 0.15
+shape <- c(2976.3, 2.31)
+rate <- c(2977.3, 1.31)
+J <- 7
+pi0 <- 0.95
+n.sim.lambda <- 10^4
+n.sim.prior <- 10^5
+n.sim.post <- 10^5
+burnin.prop <- 0.1
+theta.init <- vector(mode="numeric", length=0)
+
+n_sim <- n.sim.lambda
 
 
 
-
-
-
-
+blah(n.sim.prior, pattern, theta.init, overlap, cluster_coords, p_moves, J, prior_z, lambda)
 
 
 

### 2016-08-20 README.Rmd
Updated README

@@
 
-```{r, message=FALSE, warning=FALSE}
+```{r, message=FALSE, warning=FALSE, fig.height=8}
 plotmap(cases/population, sp.obj, log=TRUE, nclr=5)
@@
 # Compute output
-# output <- bayes_cluster(y, E, population, sp.obj, centroids, max.prop,
-#                         shape, rate, J, pi0, n.sim.lambda, n.sim.prior,
-#                         n.sim.post)
+output <- bayes_cluster(y, E, population, sp.obj, centroids, max.prop,
+                        shape, rate, J, pi0, n.sim.lambda, n.sim.prior,
+                        n.sim.post)
 ```
 
-```{r, cache=TRUE}
-plotmap(output$post.map$high.area, sp.obj)
+```{r, cache=TRUE, fig.height=8}
+plotmap(output$post_map$high_area, sp.obj)
 ```

### 2016-08-20 README.Rmd
Updated README

@@
 
-```{r, message=FALSE, warning=FALSE, fig.height=8}
+```{r, message=FALSE, warning=FALSE, fig.height=6.5}
 plotmap(cases/population, sp.obj, log=TRUE, nclr=5)
@@
 
-```{r, cache=TRUE, fig.height=8}
+```{r, cache=TRUE, fig.height=6.5}
 plotmap(output$post_map$high_area, sp.obj)

### 2018-06-02 testing/Vignettes/Install Geo Libraries/install_geo_libraries.R
Submitted to CRAN

@@ -1,0 +1,4 @@
+## ---- eval=FALSE---------------------------------------------------------
+#  library(rgdal)
+#  library(rgeos)
+

### 2018-06-02 testing/Vignettes/Old Vignette/.R

@@ -1,0 +1,332 @@
+###################################################
+### chunk number 1: 
+###################################################
+library(SpatialEpi)
+Rcode <- system.file("doc","SpatialEpi.Rnw",package="SpatialEpi")
+options(device.ask.default=FALSE)
+Stangle(Rcode)
+
+
+###################################################
+### chunk number 2: 
+###################################################
+data(scotland)
+
+
+###################################################
+### chunk number 3: 
+###################################################
+data(scotland)
+polygon <- scotland$polygon$polygon; nrepeats <- scotland$polygon$nrepeats
+names <- scotland$data$county.names
+
+spatial.polygon <- polygon2spatial.polygon(polygon,coordinate.system='+proj=utm',names,nrepeats)
+
+par(mfrow=c(1,2))
+plot(polygon,type='n',xlab="Eastings (km)",ylab="Northings (km)",main="Polygon File")
+polygon(polygon)
+
+plot(spatial.polygon,axes=TRUE)
+title(xlab="Eastings (km)",ylab="Northings (km)",main="Spatial Polygon")
+plot(spatial.polygon[23],add=TRUE,col="red")
+
+
+###################################################
+### chunk number 4: 
+###################################################
+par(mfrow=c(1,2))
+plot(polygon,type='n',xlab="Eastings (km)",ylab="Northings (km)",main="Polygon File")
+polygon(polygon)
+
+plot(spatial.polygon,axes=TRUE)
+title(xlab="Eastings (km)",ylab="Northings (km)",main="Spatial Polygon")
+plot(spatial.polygon[23],add=TRUE,col="red")
+
+
+###################################################
+### chunk number 5: 
+###################################################
+library(maps)
+county.map <- map('county',c('pennsylvania','vermont'),fill=TRUE,plot=FALSE)
+county.names <- as.character(county.map$names)
+county <- map2SpatialPolygons(county.map,IDs=county.names,proj4string=CRS("+proj=longlat"))
+state.map <- map('state',c(),fill=TRUE,plot=FALSE)
+state.names <- as.character(state.map$names)
+state <- map2SpatialPolygons(state.map,IDs=state.names,proj4string=CRS("+proj=longlat"))
+plot(county,axes=TRUE,border="red")
+plot(state,add=TRUE,lwd=2)
+
+
+###################################################
+### chunk number 6: 
+###################################################
+plot(county,axes=TRUE,border="red")
+plot(state,add=TRUE,lwd=2)
+
+
+###################################################
+### chunk number 7: 
[…]

### 2018-06-02 testing/new_kulldorff.R

@@ -1,0 +1,206 @@
+#' Generate cluster reports
+#'
+#' Return a report for a given cluster containing the same information as SatScan
+#'
+#' @inheritParams define_single_zones
+#' @inheritParams return_single_zone_areas
+#' @inheritParams kulldorff
+#' @param all_log_lkhd vector of the log likelihood of each single zone
+#' @param lambdas vector of simulated max log likelihood
+#'
+#' @return For a particular clusters a report including \describe{ 
+#'   \item{\code{location_IDs_included}}{ID's of areas in cluster, in order of distance} 
+#'   \item{\code{population}}{population of cluster} 
+#'   \item{\code{number_of_cases}}{number of cases in cluster}
+#'   \item{\code{expected_cases}}{expected number of cases in cluster}
+#'   \item{\code{SMR}}{estimated SMR of cluster}
+#'   \item{\code{log_lkhd_ratio}}{log-likelihood of cluster}
+#'   \item{\code{monte_carlo_rank}}{rank of lkhd of cluster within Monte Carlo simulated values}
+#'   \item{\code{p_value}}{Monte Carlo \eqn{p}-value}
+#'   }
+#' 
+#' @export
+#'
+#' @examples
+#' 1+1
+get_cluster_report <- function(single_zone_index, zone_info, cases, population,  
+                               expected_cases, all_log_lkhd, lambdas){
+  # Get areas included in single zone
+  cluster_areas <- return_single_zone_areas(single_zone_index, zone_info)
+  
+  # Return information
+  output <- list(
+    location_IDs_included = cluster_areas,
+    population = sum(population[cluster_areas]),	
+    number_of_cases = sum(cases[cluster_areas]),
+    expected_cases = sum(expected_cases[cluster_areas]),
+    SMR = sum(cases[cluster_areas])/sum(expected_cases[cluster_areas]),
+    log_lkhd_ratio = all_log_lkhd[single_zone_index], 
+    monte_carlo_rank = sum(lambdas >= all_log_lkhd[single_zone_index]),
+    p_value = 1-mean(lambdas < all_log_lkhd[single_zone_index])
+  )
+}
+
+
+
+#' Kulldorff method for cluster detection
+#'
+#' @inheritParams define_single_zones
+#' @param cases Vector of number of cases for each area
+#' @param expected_cases Vector of number of expected cases for each area
+#' @param alpha_level alpha-level threshold used to declare significance
[…]

### 2018-06-02 testing/roxygen/R/RcppExports.R

@@ -1,0 +1,71 @@
+# Generated by using Rcpp::compileAttributes() -> do not edit by hand
+# Generator token: 10BE3573-1514-4C36-9D1C-5A225CD40393
+
+normalize <- function(x) {
+    .Call('_SpatialEpi_normalize', PACKAGE = 'SpatialEpi', x)
+}
+
+NumericVectorEquality <- function(x, y) {
+    .Call('_SpatialEpi_NumericVectorEquality', PACKAGE = 'SpatialEpi', x, y)
+}
+
+ldnbinom <- function(y, E, a, b) {
+    .Call('_SpatialEpi_ldnbinom', PACKAGE = 'SpatialEpi', y, E, a, b)
+}
+
+ldmultinom <- function(x, prob) {
+    .Call('_SpatialEpi_ldmultinom', PACKAGE = 'SpatialEpi', x, prob)
+}
+
+coeff <- function(y_vector, E_vector, a_values, b_values, cluster_list) {
+    .Call('_SpatialEpi_coeff', PACKAGE = 'SpatialEpi', y_vector, E_vector, a_values, b_values, cluster_list)
+}
+
+ProbSampleReplace <- function(prob) {
+    .Call('_SpatialEpi_ProbSampleReplace', PACKAGE = 'SpatialEpi', prob)
+}
+
+check_overlap <- function(config, overlap) {
+    .Call('_SpatialEpi_check_overlap', PACKAGE = 'SpatialEpi', config, overlap)
+}
+
+clean_moves_matrix <- function(theta, moves, n_zones) {
+    .Call('_SpatialEpi_clean_moves_matrix', PACKAGE = 'SpatialEpi', theta, moves, n_zones)
+}
+
+return_death_moves <- function(theta) {
+    .Call('_SpatialEpi_return_death_moves', PACKAGE = 'SpatialEpi', theta)
+}
+
+return_birth_moves <- function(theta, overlap) {
+    .Call('_SpatialEpi_return_birth_moves', PACKAGE = 'SpatialEpi', theta, overlap)
+}
+
+return_local_moves <- function(theta, overlap, cluster_coords) {
+    .Call('_SpatialEpi_return_local_moves', PACKAGE = 'SpatialEpi', theta, overlap, cluster_coords)
+}
+
+MCMC_simulation <- function(n_sim, pattern, theta_init, overlap, cluster_coords, p_moves_orig, J, lkhd_z, lambda) {
[…]

### 2018-06-02 testing/roxygen/R/SpatialEpi.R

@@ -1,0 +1,20 @@
+#' SpatialEpi: Data and methods for spatial epidemiology.
+#'
+#' This package provides methods and Data for Spatial Epidemiology.
+#'
+#' @docType package
+#' @name SpatialEpi
+#' @useDynLib SpatialEpi
+#' @importFrom Rcpp sourceCpp
+#' @examples
+#' # Example usage
+#' library(SpatialEpi)
+#' data(scotland)
+#' map <- scotland$spatial.polygon
+#' y <- scotland$data$cases
+#' E <- scotland$data$expected
+#' SMR <- y/E
+#' # Plot SMR
+#' plotmap(SMR, map, nclr=9, location="topleft") 
+NULL
+#> NULL

### 2018-06-02 testing/roxygen/R/bayes_cluster.R

@@ -1,0 +1,393 @@
+#' Bayesian Cluster Detection Method
+#' 
+#' Implementation of the Bayesian Cluster detection model of Wakefield and Kim 
+#' (2013) for a study region with \code{n} areas. The prior and posterior 
+#' probabilities of each of the \code{n.zones} single zones being a 
+#' cluster/anti-cluster are estimated using Markov chain Monte Carlo. Furthermore,
+#' the posterior probability of k clusters/anti-clusters is computed.  
+#'
+#' @param y vector of length \code{n} of the observed number of disease in each area
+#' @param E vector of length \code{n} of the expected number of disease in each area
+#' @param population vector of length \code{n} of the population in each area
+#' @param sp.obj an object of class SpatialPolygons (See 
+#' \link[sp]{SpatialPolygons-class}) representing the study region
+#' @param centroids \code{n x 2} table of the (x,y)-coordinates of the area 
+#' centroids.  The coordinate system must be grid-based
+#' @param max.prop maximum proportion of the study region's population each 
+#' single zone can contain
+#' @param shape vector of length 2 of narrow/wide shape parameter for gamma 
+#' prior on relative risk
+#' @param rate vector of length 2 of narrow/wide rate parameter for gamma prior 
+#' on relative risk
+#' @param J maximum number of clusters/anti-clusters
+#' @param pi0 prior probability of no clusters/anti-clusters
+#' @param n_sim_lambda number of importance sampling iterations to estimate 
+#' lambda
+#' @param n_sim_prior number of MCMC iterations to estimate prior probabilities 
+#' associated with each single zone
+#' @param n_sim_post number of MCMC iterations to estimate posterior 
+#' probabilities associated with each single zon
+#' @param burnin.prop proportion of MCMC samples to use as burn-in. Defaults to 
+#' 10 percent
+#' @param theta.init Initial configuration used for MCMC sampling. Defaults to 
+#' \code{NULL}
+#' 
+#' @importFrom magrittr %>%
+#'
+#' @return List containing
+#' \item{prior.map}{A list containing, for each area: 1) \code{high.area} the 
+#' prior probability of cluster membership, 2) \code{low.area} anti-cluster 
[…]

### 2018-06-02 testing/roxygen/R/geographic_objects.R

@@ -1,0 +1,604 @@
+# R CMD CHECK returns NOTE that there is "no visible binding 
+# for global variable" for variables used by dplyr/ggplot2, hence this hack fix.
+utils::globalVariables(c("distance", "prop_pop", "cum_prop", "."))
+
+
+#' Define a study region's single zones.
+#' 
+#' Using the centroids of the n areas in a study region, define the set of 
+#' single zones based on the upper bound \code{pop_upper_bound} of the 
+#' proportion of the study region's population each single zone can contain.
+#' 
+#' @param centroids data frame with 2 columns of the centroids of the n areas 
+#' in the study region.
+#' @param population vector of length n of the corresponding population 
+#' counts.
+#' @param pop_upper_bound upper bound of the proportion of the study 
+#' region's population the single zones can contain.
+#'   
+#' @return A list with two objects
+#' \item{nearest_neighbors}{A list of length n of the single zones for each area} 
+#' \item{cluster_coords}{A data frame with 2 columns of the centering 
+#' and radial area for each single zone.}
+#' @references Kulldorff, M. (1997) A spatial scan statistic. 
+#' \emph{Communications in Statistics: Theory and Methods}, \bold{26}, 1481--1496.
+#' @seealso \code{\link{create_geo_objects}}
+#' @importFrom stats dist
+#' @export
+#' @importFrom magrittr %>%
+#'   
+#' @examples 
+#' data(NYleukemia)
+#' centroids <- sp::coordinates(NYleukemia$spatial.polygon)
+#' single_zones <- define_single_zones(centroids, NYleukemia$data$population, 0.15)
+define_single_zones <- function(centroids, population, pop_upper_bound) {
+  # Number of areas
+  n <- nrow(centroids)
+  # Interpoint distance matrix
+  dist_matrix <- 
+    centroids %>% 
+    dist(upper = TRUE, diag = TRUE) %>% 
+    as.matrix()
+  
+  # Area indices and proportion of the population each area contains
+  areas <- data.frame(
+    area = 1:n,
+    prop_pop= population/sum(population)
+  )
+  
+  # For each area, we compute the list of neighbors in order of distance such
+  # that the proportion of the study population that is included is less than or
+  # equal to pop_upper_bound. Save the results in a list
[…]

### 2018-06-02 testing/roxygen/R/mapping.R

@@ -1,0 +1,205 @@
+#' Plot Levels of a Variable in a Colour-Coded Map
+#' 
+#' Plot levels of a variable in a colour-coded map along with a legend.
+#'
+#' @param y variable to plot
+#' @param spatial.polygon an object of class SpatialPolygons (See 
+#' \link[sp]{SpatialPolygons-class})
+#' @param ncut number of cuts in colour levels to plot
+#' @param nlevels number of levels to include in legend
+#' @param lower lower bound of levels
+#' @param upper upper bound of levels
+#' @param main an overall title for the plot
+#' @param xlab a title for the x axis
+#' @param ylab a title for the y axis
+#'
+#' @return A choropleth map of \code{y}
+#' @export
+#' @author Jon Wakefield
+#' @author Nicky Best
+#' @author Sebastien Haneuse
+#' @author Albert Y. Kim
+#' @references Bivand, R. S., Pebesma E. J., and Gomez-Rubio V. (2008) 
+#' \emph{Applied Spatial Data Analysis with R}.  Springer Series in Statistics.
+#' @references E. J. Pebesma and R. S. Bivand. (2005) Classes and methods for 
+#' spatial data in R. \emph{R News}, \bold{5}, 9--13.  
+#' @seealso \code{\link{plotmap}}
+#' @import grDevices
+#' @import graphics
+#' @examples
+#' data(scotland)
+#' map <- scotland$spatial.polygon
+#' y <- scotland$data$cases
+#' E <- scotland$data$expected
+#' SMR <- y/E
+#' mapvariable(SMR,map,main="Scotland",xlab="Eastings (km)",ylab="Northings (km)")
+mapvariable <- function(y, spatial.polygon, ncut=1000, nlevels=10, lower=NULL, 
+                        upper=NULL, main=NULL, xlab=NULL, ylab=NULL){
+  #-------------------------------------------------------------------------------
+  # Create id indicators for coloring scheme
+  #-------------------------------------------------------------------------------
+  if (is.null(lower)) 
+    lower <- min(y) #- 0.0001
+  if (is.null(upper)) 
+    upper <- max(y) #+ 0.0001
+  
+  id <- cut(y, breaks=seq(from=lower, to=upper, length=(ncut+1)))
+  id <- as.numeric(id)
[…]

### 2018-06-02 testing/roxygen/R/other_bayes.R

@@ -1,0 +1,135 @@
+#' Empirical Bayes Estimates of Relative Risk
+#'
+#' The computes empirical Bayes estimates of relative risk of study region with 
+#' \code{n} areas, given observed and expected numbers of counts of disease and 
+#' covariate information.  
+#'
+#' @param Y a length \code{n} vector of observed cases
+#' @param E a length \code{n} vector of expected number of cases
+#' @param Xmat \code{n x p} dimension matrix of covariates
+#'
+#' @return A list with 5 elements:
+#' \item{RR}{the ecological relative risk posterior mean estimates}
+#' \item{RRmed}{the ecological relative risk posterior mean estimates}
+#' \item{beta}{the MLE's of the regression coefficients}
+#' \item{alpha}{the MLE of negative binomial dispersion parameter}
+#' \item{SMR}{the standardized mortality/morbidity ratio Y/E}
+#' @export
+#' @seealso \code{\link{scotland}}
+#' @seealso \code{\link{mapvariable}}
+#' @references Clayton D. and Kaldor J. (1987) Empirical Bayes estimates of 
+#' age-standardized relative risks for use in disease mapping.\emph{Biometrics},
+#' \bold{43}, 671--681
+#' @examples
+#' data(scotland)
+#' data <- scotland$data
+#' 
+#' x <- data$AFF
+#' Xmat <- cbind(x,x^2)
+#' results <- eBayes(data$cases,data$expected,Xmat)
+#' scotland.map <- scotland$spatial.polygon
+#' mapvariable(results$RR, scotland.map)
+eBayes <-
+  function(Y, E, Xmat=NULL){
+    # Check for covariates
+    if (is.null(Xmat)) {
+      mod <- MASS::glm.nb(Y ~ 1+offset(log(E)), link=log)
+    }else{
+      mod <- MASS::glm.nb(Y ~ Xmat+offset(log(E)), link=log)
+    }
+    
+    # Get MLE's of parameters
+    alpha <- mod$theta
+    muhat <- mod$fitted/E
+    
+    # Generate weights between global and local SMR
+    wgt <- E*muhat/(alpha+E*muhat); SMR <- Y/E
+    RR <- as.numeric(wgt*SMR + (1-wgt)*muhat)
+    RRmed <- qgamma(0.5,alpha+Y,(alpha+E*muhat)/muhat)
+    
+    # Output results
[…]

### 2018-06-02 testing/roxygen/R/other_cluster.R

@@ -1,0 +1,403 @@
+#' Besag-Newell Cluster Detection Method
+#' 
+#' Besag-Newell cluster detection method.  There are differences with the 
+#' original paper and our implementation:
+#'
+#' @param geo an \code{n x 2} table of the (x,y)-coordinates of the area centroids
+#' @param population aggregated case counts for all \code{n} areas
+#' @param cases aggregated population counts for all \code{n} areas
+#' @param expected.cases expected numbers of disease for all \code{n} areas
+#' @param k number of cases to consider
+#' @param alpha.level alpha-level threshold used to declare significance
+#'
+#' @details For the \code{population} and \code{cases} tables, the rows are 
+#' bunched by areas first, and then for each area, the counts for each strata 
+#' are listed.  It is important that the tables are balanced:  the strata 
+#' information are in the same order for each area, and counts for each 
+#' area/strata combination appear exactly once (even if zero
+#' @return List containing
+#' \item{clusters}{information on all clusters that are \eqn{\alpha}-level 
+#' significant, in decreasing order of the \eqn{p}-value}
+#' \item{p.values}{for each of the \eqn{n} areas, \eqn{p}-values of each cluster 
+#' of size at least \eqn{k}}
+#' \item{m.values}{for each of the \eqn{n} areas, the number of areas need to 
+#' observe at least \eqn{k} cases}
+#' \item{observed.k.values}{based on \code{m.values}, the actual number of cases 
+#' used to compute the \eqn{p}-values}
+#' @export
+#' @seealso \code{\link{pennLC}}
+#' @seealso \code{\link{expected}}
+#' @references Besag J. and Newell J. (1991) The Detection of Clusters in Rare 
+#' Diseases \emph{Journal of the Royal Statistical Society. Series A (Statistics 
+#' in Society)}, \bold{154}, 143--155
+#'
+#' @examples
+#' # Load Pennsylvania Lung Cancer Data
+#' data(pennLC)
+#' data <- pennLC$data
+#' 
+#' # Process geographical information and convert to grid
+#' geo <- pennLC$geo[,2:3]
+#' geo <- latlong2grid(geo)
+#' 
+#' # Get aggregated counts of population and cases for each county
+#' population <- tapply(data$population,data$county,sum)
+#' cases <- tapply(data$cases,data$county,sum)
+#' 
+#' # Based on the 16 strata levels, computed expected numbers of disease
+#' n.strata <- 16
+#' expected.cases <- expected(data$population, data$cases, n.strata)
+#' 
[…]

### 2018-06-02 testing/testing.R

@@ -1,0 +1,241 @@
+# Testing bayes_cluster
+data(NYleukemia)
+sp.obj <- NYleukemia$spatial.polygon
+centroids <- latlong2grid(NYleukemia$geo[, 2:3])
+population <- NYleukemia$data$population
+cases <- NYleukemia$data$cases
+y <- cases
+E <- expected(population, cases, 1)
+max.prop <- 0.15
+shape <- c(2976.3, 2.31)
+rate <- c(2977.3, 1.31)
+J <- 7
+pi0 <- 0.95
+n.sim.lambda <- 10^4
+n.sim.prior <- 10^5
+n.sim.post <- 10^5
+burnin.prop <- 0.1
+theta.init <- vector(mode="numeric", length=0)
+
+n_sim <- n.sim.lambda
+
+
+
+blah(n.sim.prior, pattern, theta.init, overlap, cluster_coords, p_moves, J, prior_z, lambda)
+
+
+
+# R CMD CHECK returns NOTE that there is "no visible binding
+# for global variable" for variables used by dplyr/ggplot2, hence this hack fix.
+# utils::globalVariables(c("lat", "long", "group", ".", "variable"))
+
+#' Plot ggplot compliant map
+#'
+#' As described in
+#' \url{https://github.com/hadley/ggplot2/wiki/plotting-polygon-shapefiles}
+#'
+#' @param sp_obj SpatialPolygonsDataFrame object
+#' @param variable_name Variable within data frame \code{sp_obj@@data}
+#' @param type string indicating type of plot. Must be either
+#'   \code{"continuous"} for continuous scale, \code{"discrete"} for categorical
+#'   variables, \code{"quantile"} for quantile-based bins, or
+#'   \code{"equidistant"} for equally-sized bins.
+#' @param n_int For non-continuous scales, number of bins.
+#' @param x_label Label for eastings
+#' @param lwd line width between areas
+#' @param y_label Label for northings
+#' @param title Plot title
+#' @param legend_title label for legend. Defaults to \code{variable_name}.
+#'
+#' @import maptools
+#'
+#' @return a ggplot object
+#' @export
+map_variable <- function(sp_obj, variable_name = NULL, type = "continuous", n_int=5,
+                         x_label = "eastings", y_label = "northings", lwd=0.5,
[…]

### 2018-06-02 README.Rmd
About to release on CRAN
Update README

@@
 
-We run the Bayesian Cluster Detection method from Wakefield and Kim (2013)
+We run the Bayesian Cluster Detection method from [Wakefield and Kim (2013)](https://www.researchgate.net/publication/235896508_A_Bayesian_model_for_cluster_detection):
 
@@
 
-```{r, cache=TRUE, fig.height=6.5}
+```{r, fig.height=6.5}
 plotmap(output$post_map$high_area, sp.obj)

### 2021-06-10 tests/testthat.R
refactoring expected function

@@ -1,0 +1,9 @@
+
+library(testthat)
+library(SpatialEpi)
+
+test_that("population_test", {
+  
+ 
+})
+

### 2021-06-10 R/expected.R
revert test

@@ -1,12 +1,6 @@
 expected <- 
   function(population, cases, n.strata){
     
-  if(!is.data.frame(population))
-    stop()
-  
-  if(!is.data.frame(cases))
-    stop()
-  
   
 
 n <- length(population)/n.strata

### 2021-06-10 tests/testthat.R

@@ -1,9 +1,4 @@
-
 library(testthat)
 library(SpatialEpi)
 
-test_that("population_test", {
-  
- 
-})
-
+test_check("SpatialEpi")

### 2021-06-10 README.Rmd
refactoring expected function
Trivial change

@@
 
-[![Travis-CI Build Status](https://travis-ci.org/rudeboybert/SpatialEpi.svg?branch=master)](https://travis-ci.org/rudeboybert/SpatialEpi)
-[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/SpatialEpi)](http://cran.r-project.org/package=SpatialEpi)
-[![CRAN RStudio mirror downloads](http://cranlogs.r-pkg.org/badges/SpatialEpi)](http://www.r-pkg.org/pkg/SpatialEpi)
+[![R-CMD-check](https://github.com/rudeboybert/SpatialEpi/workflows/R-CMD-check/badge.svg)](https://github.com/rudeboybert/SpatialEpi/actions)
+[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/SpatialEpi)](http://cran.r-project.org/package=SpatialEpi) [![CRAN RStudio mirror downloads](http://cranlogs.r-pkg.org/badges/SpatialEpi)](http://www.r-pkg.org/pkg/SpatialEpi)
 
@@
 
+Note: You may need to install the `cpp11` for all C++ code to compile correctly.
+
 ## Example
 
-We load the data and convert the coordinate system from latitude/longitude to a 
-grid-based system.
+We load the data and convert the coordinate system from latitude/longitude to a grid-based system.
 
@@
 ```{r, fig.height=6.5}
-plotmap(output$post_map$high_area, sp.obj)
+plotmap(output$post.map$high.area, sp.obj)
 ```

### 2021-06-12 R/PennLC.R
working on converting sp to sf committing so I can save my work

@@ -1,0 +1,5 @@
+library(SpatialEpi)
+library(sf)
+
+
+st_as_sf(pennLC$spatial.polygon)

### 2021-07-12 R/LogNormalPriorCh.R
commtting to save

@@ -17,9 +17,10 @@
 #' \item{sigma}{variance of log-normal distribution}
 #' @export
 #'
-#' @examples # Calibrate the log-normal distribution s.t. the 95% confidence interval is [0.2, 5]
+#' @examples 
+#' #Calibrate the log-normal distribution s.t. the 95% confidence interval is [0.2, 5]
 #' param <- LogNormalPriorCh(0.2, 5, 0.025, 0.975)
-#' curve(dlnorm(x,param$mu,param$sigma), from=0, to=6, ylab="density"
+#' curve(dlnorm(x,param$mu,param$sigma), from=0, to=6, ylab="density
 LogNormalPriorCh <-
 function(theta1, theta2, prob1, prob2){	
 zq1 <- qnorm(prob1)

### 2021-07-26 README.Rmd
Fixed all url's

@@
 [![R-CMD-check](https://github.com/rudeboybert/SpatialEpi/workflows/R-CMD-check/badge.svg)](https://github.com/rudeboybert/SpatialEpi/actions)
-[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/SpatialEpi)](http://cran.r-project.org/package=SpatialEpi) [![CRAN RStudio mirror downloads](http://cranlogs.r-pkg.org/badges/SpatialEpi)](http://www.r-pkg.org/pkg/SpatialEpi)
+[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/SpatialEpi)](https://cran.r-project.org/package=SpatialEpi) [![CRAN RStudio mirror downloads](https://cranlogs.r-pkg.org/badges/SpatialEpi)](https://www.r-pkg.org/pkg/SpatialEpi)
 

### 2021-07-26 R/pennLC.R

@@ -17,7 +17,7 @@
 #'   
 #' }
 #' 
-#' @source Population data was obtained from the 2000 decennial census, lung cancer and smoking data were obtained from the Pennsylvania Department of Health website:  \url{http://www.dsf.health.state.pa.us/}
+#' @source Population data was obtained from the 2000 decennial census, lung cancer and smoking data were obtained from the Pennsylvania Department of Health website:  \url{https://www.health.pa.gov/Pages/default.aspx}
 #' 
 #' 
 #' @examples 

### 2021-07-26 R/sf_objects.R

@@ -17,7 +17,7 @@
 #'  \item{smoking}{Overall county smoking rate (not broken down by strata)}
 #'  \item{geometry}{Geometric representation of counties in Pennsylvania}
 #' }
-#' @source Population data was obtained from the 2000 decennial census, lung cancer and smoking data were obtained from the Pennsylvania Department of Health website:\url{http://www.dsf.health.state.pa.us/}.
+#' @source Population data was obtained from the 2000 decennial census, lung cancer and smoking data were obtained from the Pennsylvania Department of Health website:\url{https://www.health.pa.gov/Pages/default.aspx}.
 #' @examples 
 #' library(ggplot2)
 #' library(dplyr)

### 2021-07-26 src/RcppExports.cpp

@@ -6,6 +6,11 @@
 
 using namespace Rcpp;
 
+#ifdef RCPP_USE_GLOBAL_ROSTREAM
+Rcpp::Rostream<true>&  Rcpp::Rcout = Rcpp::Rcpp_cout_get();
+Rcpp::Rostream<false>& Rcpp::Rcerr = Rcpp::Rcpp_cerr_get();
+#endif
+
 // normalize
 NumericVector normalize(NumericVector x);
 RcppExport SEXP _SpatialEpi_normalize(SEXP xSEXP) {

### 2021-07-27 R/EBpostthresh.R
Ran roxygen2md::roxygen2md()

@@ -8,7 +8,7 @@ globalVariables(c(
 #'
 #' @description This function produces the posterior probabilities of exceeding a threshold given a gamma distributions with parameters (alpha+Y, (alpha+E*mu)/mu) where mu = exp(x beta). This model arises from Y being Poisson with mean theta times E where theta is the relative risk and E are the expected numbers. The prior on theta is gamma with parameters alpha and beta. The parameters alpha and beta may be estimated using empirical Bayes.
 #' @author Jon Wakefield
-#' @seealso  \code{\link{eBayes}}
+#' @seealso  [eBayes()]
 #' 
 #' @param Y observed disease counts
 #' @param E expected disease counts

### 2021-07-27 R/NYleukemia.R

@@ -1,10 +1,10 @@
 #' Upstate New York Leukemia Data
 #' 
 #' 
-#' Census tract level (\code{n=281}) leukemia data for the 8 counties in upstate New York from 1978-1982, paired with population data from the 1980 census.  
+#' Census tract level (`n=281`) leukemia data for the 8 counties in upstate New York from 1978-1982, paired with population data from the 1980 census.  
 #' Note that 4 census tracts were completely surrounded by another unique census tract; 
-#' when applying the Bayesian cluster detection model in \code{\link{bayes_cluster}},
-#' we merge them with the surrounding census tracts yielding \code{n=277} areas.
+#' when applying the Bayesian cluster detection model in [bayes_cluster()],
+#' we merge them with the surrounding census tracts yielding `n=277` areas.
 #' 
 #' 
 #' @format 
@@ -18,7 +18,7 @@
 #'   
 #' }
 #' 
-#' @references  Turnbull, B. W. et al (1990) Monitoring for clusters of disease: application to leukemia incidence in upstate New York \emph{American Journal of Epidemiology}, \bold{132}, 136--143
+#' @references  Turnbull, B. W. et al (1990) Monitoring for clusters of disease: application to leukemia incidence in upstate New York *American Journal of Epidemiology*, **132**, 136--143
 #' 
 #' @examples 
 #' ## Load data and convert coordinate system from latitude/longitude to grid

### 2021-07-27 R/NYleukemia_sf.R

@@ -1,10 +1,10 @@
 #' Upstate New York Leukemia 
 #' 
 #' 
-#' Census tract level (\code{n=281}) leukemia data for the 8 counties in upstate New York from 1978-1982, paired with population data from the 1980 census.  
+#' Census tract level (`n=281`) leukemia data for the 8 counties in upstate New York from 1978-1982, paired with population data from the 1980 census.  
 #' Note that 4 census tracts were completely surrounded by another unique census tract; 
-#' when applying the Bayesian cluster detection model in \code{\link{bayes_cluster}},
-#' we merge them with the surrounding census tracts yielding \code{n=277} areas.
+#' when applying the Bayesian cluster detection model in [bayes_cluster()],
+#' we merge them with the surrounding census tracts yielding `n=277` areas.
 #' 
 #' 
 #' @format An sf 'POLYGON' data frame with 281 rows  and 4 variables:
@@ -15,7 +15,7 @@
 #'   \item{censustract.FIPS}{11-digit Federal Information Processing System identification number for each county}
 #'   
 #' }
-#' @source Turnbull, B. W. et al (1990) Monitoring for clusters of disease: application to leukemia incidence in upstate New York \emph{American Journal of Epidemiology}, \bold{132}, 136--143
+#' @source Turnbull, B. W. et al (1990) Monitoring for clusters of disease: application to leukemia incidence in upstate New York *American Journal of Epidemiology*, **132**, 136--143
 #' 
 #' @examples 
 #' 

### 2021-07-27 R/besag_newell.R

@@ -2,18 +2,18 @@
 #' 
 #' @description Besag-Newell cluster detection method.  There are differences with the original paper and our implementation:
 #' \itemize{
-#' \item{we base our analysis on \eqn{k} cases, rather than \eqn{k} \emph{other} cases as prescribed in the paper.} 
-#' \item we do not subtract 1 from the \emph{accumulated numbers of other cases} and \emph{accumulated numbers of others at risk}, as was prescribed in the paper to discount selection bias
+#' \item{we base our analysis on \eqn{k} cases, rather than \eqn{k} *other* cases as prescribed in the paper.} 
+#' \item we do not subtract 1 from the *accumulated numbers of other cases* and *accumulated numbers of others at risk*, as was prescribed in the paper to discount selection bias
 #' \item M is the total number of areas included, not the number of additional areas included.  i.e. \eqn{M} starts at 1, not 0.
 #' \item p-values are not based on the original value of \eqn{k}, rather the actual number of cases observed until we view \eqn{k} or more cases.  Ex:  if \eqn{k = 10}, but as we consider neighbors we encounter 1, 2, 9 then 12 cases, we base our \eqn{p}-values on \eqn{k=12}
 #' \item we do not provide a Monte-Carlo simulated \eqn{R}:  the number of tests that attain significance at a fixed level \eqn{\alpha}
 #' }
 #' The first two and last differences are because we view the testing on an area-by-area level, rather than a case-by-case level.
 #' 
-#' @param geo an \code{n x 2} table of the (x,y)-coordinates of the area centroids
-#' @param population aggregated population counts for all \code{n} areas
-#' @param cases aggregated case counts for all \code{n} areas
-#' @param expected.cases expected numbers of disease for all \code{n} areas
+#' @param geo an `n x 2` table of the (x,y)-coordinates of the area centroids
+#' @param population aggregated population counts for all `n` areas
+#' @param cases aggregated case counts for all `n` areas
+#' @param expected.cases expected numbers of disease for all `n` areas
 #' @param k number of cases to consider
 #' @param alpha.level alpha-level threshold used to declare significance
 #'
@@ -23,9 +23,9 @@
 #'
 #'
 #' @details 
[…]

### 2021-07-27 R/circle.R

@@ -11,9 +11,9 @@ globalVariables(c(
 #' 
 #' @author Albert Y. Kim
 #' 
-#' @param geo A \code{n x 2} table of the x-coordinate and y-coordinates of the centroids of each area
-#' @param cluster.center The area index (an integer between \code{1} and \code{n}) indicating the center of the circle
-#' @param cluster.end The area index (an integer between \code{1} and \code{n}) indicating the area at the end of the circle
+#' @param geo A `n x 2` table of the x-coordinate and y-coordinates of the centroids of each area
+#' @param cluster.center The area index (an integer between `1` and `n`) indicating the center of the circle
+#' @param cluster.end The area index (an integer between `1` and `n`) indicating the area at the end of the circle
 #'
 #' @return
 #'  \item{cluster.radius}{A data frame that you can plot}

### 2021-07-27 R/eBayes.R

@@ -6,13 +6,13 @@ globalVariables(c(
 
 
 #' Empirical Bayes Estimates of Relative Risk
-#' @description The computes empirical Bayes estimates of relative risk of study region with \code{n} areas, given observed and expected numbers of counts of disease and covariate information.
+#' @description The computes empirical Bayes estimates of relative risk of study region with `n` areas, given observed and expected numbers of counts of disease and covariate information.
 #' 
-#' @param Y a length \code{n} vector of observed cases
-#' @param E a length \code{n} vector of expected number of cases
-#' @param Xmat \code{n x p} dimension matrix of covariates
+#' @param Y a length `n` vector of observed cases
+#' @param E a length `n` vector of expected number of cases
+#' @param Xmat `n x p` dimension matrix of covariates
 #'
-#' @references Clayton D. and Kaldor J. (1987) Empirical Bayes estimates of age-standardized relative risks for use in disease mapping.  \emph{Biometrics}, \bold{43}, 671--681
+#' @references Clayton D. and Kaldor J. (1987) Empirical Bayes estimates of age-standardized relative risks for use in disease mapping.  *Biometrics*, **43**, 671--681
 #' 
 #' 
 #' @return

### 2021-07-27 R/expected.R

@@ -5,9 +5,9 @@
 #' @param cases a vector of the corresponding number of cases
 #' @param n.strata number of strata considered
 #' 
-#' @details The \code{population} and \code{cases} vectors must be \emph{balanced}: all counts are sorted by area first, and then within each area the counts for all strata are listed (even if 0 count) in the same order.
+#' @details The `population` and `cases` vectors must be *balanced*: all counts are sorted by area first, and then within each area the counts for all strata are listed (even if 0 count) in the same order.
 #'
-#' @references Elliot, P. et al. (2000) \emph{Spatial Epidemiology:  Methods and Applications}.  Oxford Medical Publications.
+#' @references Elliot, P. et al. (2000) *Spatial Epidemiology:  Methods and Applications*.  Oxford Medical Publications.
 #' @author Albert Y. Kim
 #' @return
 #' \item{expected.cases}{a vector of the expected numbers of disease for each area}

### 2021-07-27 R/grid2latlong.R

@@ -10,7 +10,7 @@ globalVariables(c(
 #'
 #' @description Convert geographic coordinates from Universal Transverse Mercator system to Latitude/Longitude.
 #' 
-#' @param input A data frame with columns named \code{x} and \code{y} of the UTM coordinates to convert or  an \code{n x 2} matrix of grid coordinates or an object of class SpatialPolygons (See \link[sp]{SpatialPolygons-class})
+#' @param input A data frame with columns named `x` and `y` of the UTM coordinates to convert or  an `n x 2` matrix of grid coordinates or an object of class SpatialPolygons (See [SpatialPolygons-class][sp::SpatialPolygons-class])
 #'
 #' @author Lance A. Waller
 #' 

### 2021-07-27 R/latlong2grid.R

@@ -9,7 +9,7 @@ globalVariables(c(
 #'
 #' @description Convert geographic latitude/longitude coordinates to kilometer-based grid coordinates.
 #' 
-#' @param input either an \code{n x 2} matrix of longitude and latitude coordinates in decimal format or an object of class SpatialPolygons 
+#' @param input either an `n x 2` matrix of longitude and latitude coordinates in decimal format or an object of class SpatialPolygons 
 #'
 #' @note Rough conversion of US lat/long to km (used by GeoBUGS):  (see also forum.swarthmore.edu/dr.math/problems/longandlat.html).  Radius of earth: r = 3963.34 (equatorial) or 3949.99 (polar) mi = 6378.2 or 6356.7 km, which implies: km per mile  = 1.609299 or 1.609295 a change of 1 degree of latitude corresponds to the same number of km, regardless of longitude.  arclength=r*theta, so the multiplier for coord\$y should probably be just the radius of earth.
 #' On the other hand, a change of 1 degree in longitude corresponds to a different distance, depending on latitude.  (at N pole, the change is essentially 0.  at the equator, use equatorial radius.

### 2021-07-27 R/mapvariable.R

@@ -7,7 +7,7 @@ globalVariables(c(
 #' @description Plot levels of a variable in a colour-coded map along with a legend.
 #' 
 #' @param y variable to plot
-#' @param spatial.polygon an object of class SpatialPolygons (See \link[sp]{SpatialPolygons-class})
+#' @param spatial.polygon an object of class SpatialPolygons (See [SpatialPolygons-class][sp::SpatialPolygons-class])
 #' @param ncut number of cuts in colour levels to plot
 #' @param nlevels number of levels to include in legend
 #' @param lower lower bound of levels
@@ -16,12 +16,12 @@ globalVariables(c(
 #' @param xlab a title for the x axis
 #' @param ylab a title for the y axis
 #'
-#' @references Bivand, R. S., Pebesma E. J., and Gomez-Rubio V. (2008) \emph{Applied Spatial Data Analysis with R}.  Springer Series in Statistics.
-#' E. J. Pebesma and R. S. Bivand. (2005) Classes and methods for spatial data in R. \emph{R News}, \bold{5}, 9--13. 
+#' @references Bivand, R. S., Pebesma E. J., and Gomez-Rubio V. (2008) *Applied Spatial Data Analysis with R*.  Springer Series in Statistics.
+#' E. J. Pebesma and R. S. Bivand. (2005) Classes and methods for spatial data in R. *R News*, **5**, 9--13. 
 #' 
 #' @author Jon Wakefield, Nicky Best, Sebastien Haneuse, and Albert Y. Kim
 #' 
-#' @return A map colour-coded to indicate the different levels of \code{y}
+#' @return A map colour-coded to indicate the different levels of `y`
 #' 
 #' @export
 #' @import sp

### 2021-07-27 R/polygon2spatial_polygon.R

@@ -15,14 +15,14 @@ globalVariables(c(
 #' @param area.names names of all areas
 #' @param nrepeats number of sub polygons for each area
 #'
-#' @details Just as when plotting with the \code{\link[graphics]{polygon}} function, it is assumed that each subpolygon is to be closed by joining the last point to the first point.  In the matrix \code{poly}, NA values separate complete subpolygons. 
-#' \code{coordinate.system} must be either \code{'+proj=utm'} or \code{'+proj=longlat'}.
-#' In the case with an area consists of more than one separate closed polygon, \code{nrepeats} specifies the number of closed polygons associated with each area.
+#' @details Just as when plotting with the [graphics::polygon()] function, it is assumed that each subpolygon is to be closed by joining the last point to the first point.  In the matrix `poly`, NA values separate complete subpolygons. 
+#' `coordinate.system` must be either `'+proj=utm'` or `'+proj=longlat'`.
+#' In the case with an area consists of more than one separate closed polygon, `nrepeats` specifies the number of closed polygons associated with each area.
 #' 
-#' @references Bivand, R. S., Pebesma E. J., and Gomez-Rubio V. (2008) \emph{Applied Spatial Data Analysis with R}.  Springer Series in Statistics. E. J. Pebesma and R. S. Bivand. (2005) Classes and methods for spatial data in R. \emph{R News}, \bold{5}, 9--13.  
+#' @references Bivand, R. S., Pebesma E. J., and Gomez-Rubio V. (2008) *Applied Spatial Data Analysis with R*.  Springer Series in Statistics. E. J. Pebesma and R. S. Bivand. (2005) Classes and methods for spatial data in R. *R News*, **5**, 9--13.  
 #' @author Albert Y. Kim
 #' @return
-#' An object of class SpatialPolygons (See \link[sp]{SpatialPolygons-class} from the \pkg{sp} package).
+#' An object of class SpatialPolygons (See [SpatialPolygons-class][sp::SpatialPolygons-class] from the \pkg{sp} package).
 #' @export
 #'
 #' @examples

### 2021-07-27 R/scotland.R

@@ -8,12 +8,12 @@
 #' \describe{
 #'   \item{geo}{a table of county IDs, x-coordinates (eastings) and y-coordinates (northings) of the geographic centroid of each county.}
 #'   \item{data}{a table of county IDs, number of cases, population and strata information}
-#'   \item{spatial.polygon}{a Spatial Polygons class (See \link[sp]{SpatialPolygons-class}) map of Scotland}
-#'   \item{polygon}{a polygon map of Scotland (See \code{\link{polygon2spatial_polygon}}}
+#'   \item{spatial.polygon}{a Spatial Polygons class (See [SpatialPolygons-class][sp::SpatialPolygons-class]) map of Scotland}
+#'   \item{polygon}{a polygon map of Scotland (See [polygon2spatial_polygon()]}
 #' }
-#' @source Kemp I., Boyle P., Smans M. and Muir C. (1985) Atlas of cancer in Scotland, 1975-1980, incidence and epidemiologic perspective \emph{International Agency for Research on Cancer} \bold{72}.
+#' @source Kemp I., Boyle P., Smans M. and Muir C. (1985) Atlas of cancer in Scotland, 1975-1980, incidence and epidemiologic perspective *International Agency for Research on Cancer* **72**.
 #'
-#' @references Clayton D. and Kaldor J. (1987) Empirical Bayes estimates of age-standardized relative risks for use in disease mapping.  \emph{Biometrics}, \bold{43}, 671--681.
+#' @references Clayton D. and Kaldor J. (1987) Empirical Bayes estimates of age-standardized relative risks for use in disease mapping.  *Biometrics*, **43**, 671--681.
 #' 
 #' @examples
 #' data(scotland)

### 2021-07-27 R/sf_objects.R

@@ -17,7 +17,7 @@
 #'  \item{smoking}{Overall county smoking rate (not broken down by strata)}
 #'  \item{geometry}{Geometric representation of counties in Pennsylvania}
 #' }
-#' @source Population data was obtained from the 2000 decennial census, lung cancer and smoking data were obtained from the Pennsylvania Department of Health website:\url{https://www.health.pa.gov/Pages/default.aspx}.
+#' @source Population data was obtained from the 2000 decennial census, lung cancer and smoking data were obtained from the Pennsylvania Department of Health website:<https://www.health.pa.gov/Pages/default.aspx>.
 #' @examples 
 #' library(ggplot2)
 #' library(dplyr)

### 2021-07-27 R/zones.R

@@ -1,21 +1,21 @@
 #' Create set of all single zones and output geographical information
 #'
-#' @description Based on the population counts and centroid coordinates of each of \code{n} areas, output the set of \code{n.zones} single zones as defined by Kulldorff and other geographical information.
-#' @param geo \code{n x 2} table of the (x,y)-coordinates of the area centroids
+#' @description Based on the population counts and centroid coordinates of each of `n` areas, output the set of `n.zones` single zones as defined by Kulldorff and other geographical information.
+#' @param geo `n x 2` table of the (x,y)-coordinates of the area centroids
 #' @param population a vector of population counts of each area
 #' @param pop.upper.bound maximum proportion of study region each zone can contain
 #'
-#' @references Kulldorff, M. (1997) A spatial scan statistic. \emph{Communications in Statistics: Theory and Methods}, \bold{26}, 1481--1496.
+#' @references Kulldorff, M. (1997) A spatial scan statistic. *Communications in Statistics: Theory and Methods*, **26**, 1481--1496.
 #' Kulldorff M. and Nagarwalla N. (1995) Spatial disease clusters: Detection and Inference.
-#' \emph{Statistics in Medicine}, \bold{14}, 799--810.
+#' *Statistics in Medicine*, **14**, 799--810.
 #' 
 #' @author Albert Y. Kim
 #' 
 #' @return
 #' A list containing
-#' \item{nearest.neighbors}{list of \code{n} elements, where each element is a vector of the nearest neighbors in order of distance up until \code{pop.upper.bound} of the total population is attained}
-#' \item{cluster.coords}{\code{n.zones x 2} table of the center and the radial area for each zone}
-#' \item{dist}{\code{n x n} inter-point distance matrix of the centroids}
+#' \item{nearest.neighbors}{list of `n` elements, where each element is a vector of the nearest neighbors in order of distance up until `pop.upper.bound` of the total population is attained}
+#' \item{cluster.coords}{`n.zones x 2` table of the center and the radial area for each zone}
+#' \item{dist}{`n x n` inter-point distance matrix of the centroids}
 #'
 #' 
 #' @export

### 2021-08-01 vignettes/install_geo_libraries.R
gitignore clean

@@ -1,13 +1,0 @@
-## ---- include = FALSE---------------------------------------------------------
-knitr::opts_chunk$set(
-  collapse = TRUE,
-  comment = "#>"
-)
-
-## ----setup--------------------------------------------------------------------
-library(SpatialEpi)
-
-## ---- eval=FALSE--------------------------------------------------------------
-#  library(rgdal)
-#  library(rgeos)
-

### 2021-08-01 README.Rmd
Update README

@@
 
-Note: You may need to install the `cpp11` for all C++ code to compile correctly.
+Note: In order for all C++ code to compile correctly you may need to
+
+1. Install the `cpp11` package
+1. Install an older version of `RcppArmadillo` by running  
+    ```{r, eval=FALSE}
+    packageurl <- "https://cran.r-project.org/src/contrib/Archive/RcppArmadillo/RcppArmadillo_0.9.900.3.0.tar.gz"
+install.packages(packageurl, repos=NULL, type="source")
+    ```
 

### 2021-11-12 R/grid2latlong.R
Remove all proj=utm

@@ -65,7 +65,7 @@ if(is(input)[1] == "SpatialPolygons"){
       ID=input@polygons[[i]]@ID
     )	
   }
-  output <- SpatialPolygons(input@polygons,proj4string=CRS("+proj=utm"))
+  output <- SpatialPolygons(input@polygons,proj4string=CRS("+proj=eqc +lat_ts=0 +lat_0=0 +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"))
 
   
 #-------------------------------------------------------------------------------

### 2021-11-12 R/latlong2grid.R

@@ -51,7 +51,7 @@ sine51 <- sin( 51.5*toradians )
 #-------------------------------------------------------------------------------
 # If a Spatial Polygon
 #-------------------------------------------------------------------------------
-if(is(input)[1] == "SpatialPolygons"){  
+if(is(input)[1] == "SpatialPolygons"){
   for( i in 1:length(input@polygons) ){
     # for all Polygons's in polygon
     for( j in 1:length(input@polygons[[i]]@Polygons) ){
@@ -75,7 +75,7 @@ if(is(input)[1] == "SpatialPolygons"){
     )	
   }
   
-  output <- SpatialPolygons(input@polygons,proj4string=CRS("+proj=utm"))
+  output <- SpatialPolygons(input@polygons,proj4string=CRS("+proj=eqc +lat_ts=0 +lat_0=0 +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"))
 
   
 #-------------------------------------------------------------------------------  

### 2021-11-12 R/polygon2spatial_polygon.R

@@ -16,7 +16,6 @@ globalVariables(c(
 #' @param nrepeats number of sub polygons for each area
 #'
 #' @details Just as when plotting with the [graphics::polygon()] function, it is assumed that each subpolygon is to be closed by joining the last point to the first point.  In the matrix `poly`, NA values separate complete subpolygons. 
-#' `coordinate.system` must be either `'+proj=utm'` or `'+proj=longlat'`.
 #' In the case with an area consists of more than one separate closed polygon, `nrepeats` specifies the number of closed polygons associated with each area.
 #' 
 #' @references Bivand, R. S., Pebesma E. J., and Gomez-Rubio V. (2008) *Applied Spatial Data Analysis with R*.  Springer Series in Statistics. E. J. Pebesma and R. S. Bivand. (2005) Classes and methods for spatial data in R. *R News*, **5**, 9--13.  
@@ -30,7 +29,8 @@ globalVariables(c(
 #' data(scotland)
 #' 
 #' polygon <- scotland$polygon$polygon
-#' coord.system <- '+proj=utm'
+#' coord.system <- "+proj=eqc +lat_ts=0 +lat_0=0 +lon_0=0 +x_0=0 +y_0=0 "
+#' coord.system <- paste(coord.system, "+ellps=WGS84 +datum=WGS84 +units=m +no_defs", sep = "")
 #' names <- scotland$data$county.names
 #' nrepeats <- scotland$polygon$nrepeats
 #' 
@@ -55,7 +55,7 @@ function(poly, coordinate.system, area.names = NULL, nrepeats = NULL){
 # Deal with non-specified values
 #-------------------------------------------------------------------------------
 if(missing(coordinate.system)){
-	stop("Coordinate system must be specified: '+proj=utm' or '+proj=longlat'.")
+	stop("Coordinate system must be specified")
 }
 
 if(is.null(nrepeats)){

### 2023-02-21 R/leglabs.R
Getting ready for version bump

@@ -1,0 +1,30 @@
+#' Make legend labels
+#'
+#' leglabs makes character strings from the same break points. This function was copied from the soon-to-be 
+#' deprecated `maptools` package with permission from author Roger Bivand
+#'
+#' @param vec vector of break values
+#' @param under character value for under
+#' @param over character value for over
+#' @param between character value for between
+#' @param reverse flag to reverse order of values, you will also need to reorder colours, see example
+#'
+#' @author Roger Bivand, Nick Bearman, Nicholas Lewin-Koh 
+leglabs <- function(vec, under = "under", over = "over", between = "-", 
+                    reverse = FALSE) {
+  x <- vec
+  lx <- length(x)
+  if (lx < 3) 
+    stop("vector too short")
+  if (reverse) {
+    x <- rev(x)
+    under <- "over"
+    over <- "under"
+  }
+  res <- character(lx - 1)
+  res[1] <- paste(under, x[2])
+  for (i in 2:(lx - 2)) res[i] <- paste(x[i], between, x[i + 
+                                                           1])
+  res[lx - 1] <- paste(over, x[lx - 1])
+  res
+}

### 2023-02-21 R/plotmap.R

@@ -21,7 +21,6 @@ globalVariables(c(
 #' @return
 #' A map colour-coded to indicate the different levels of `values`.
 #' @export
-#' @importFrom maptools leglabs
 #'
 #' @examples
 #' ## Load data
